---
title: "Financial Burn Down Scenarios"
author: "`r Sys.getenv('USER')`"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=3.5,
                      fig.width=6)

library(FinCal)
library(tidyverse)
library(lubridate)
library(here)
library(PerformanceAnalytics)
library(plotly)
library(scales)
library(gridExtra)
library(DT)

options(scipen = 10)
theme_set(theme_light())

```

## Intro

Exploration of some different scenarios for financial burndown based on various assumptions around present value, annual expenditures, annual returns.

## References

* 
* [Time value of money calculations in R - no pkg](https://www.codingfinance.com/post/2018-03-18-tvm/)

### Inputs & Output

Inputs needed:

* starting balance
* annual drawdown
* return on balance
* number of years of drawdown period

Output is remaining balance at end of period.

```{r, include=FALSE}
## test area for single calculation to get end balance
startbal <- -600000
draw <- 60000
return <- 0.04
yrs <- 4

remain_end <- fv(r=return, n=yrs, pv=startbal, pmt=draw, type=0)
```

## Single Scenario

```{r}
## loop through each year, new calc based on balance at end of each year to get year-by-year schedule
startbal <- -600000
draw <- 60000
return <- 0.04
yrs <- 10

start_bal <- startbal
sched_all_s <- data.frame()
for(y in 1:yrs){
  remain <- fv(r=return, n=1, pv=start_bal, pmt=draw, type=0)
  sched <- data.frame('year'=y,
                      'balance'=remain)
  sched_all_s <- bind_rows(sched_all_s, sched)
  start_bal <- remain*-1
}
```

* Start: `r paste0("$",format(startbal*-1, big.mark=",", digits=0))`
* Draw: `r paste0("$",format(draw, big.mark=",", digits=0), " / yr")`
* Return rate: `r paste0(return*100,"%")`
* Years: `r yrs`

At end of `r yrs` years, you will have `r paste0("$",format(remain, big.mark=",", digits=0))` left.

```{r}
sched_all_s %>% ggplot(aes(x=as.factor(year), y=balance))+geom_line(group=1)+
  scale_y_continuous(labels=comma)+
  labs(title=paste0("$",format(remain, big.mark=",", digits=0), " left after ", yrs, " years."),
       y="", x="year from start")
```

## Multiple Scenarios

```{r}

## loop through each year, new calc based on balance at end of each year to get year-by-year schedule
scenarios <- tribble(
  ~startbal, ~draw, ~return, ~yrs,
  -600000, 60000, 0.04, 20,
  -700000, 60000, 0.03, 20,
  -700000, 60000, 0.02, 20,
  -800000, 60000, 0.02, 20,
  -800000, 60000, 0.03, 20,
  -1000000, 70000, 0.03, 20,
  -1000000, 60000, 0.04, 20,
  -800000, 60000, 0.04, 20
)
scenarios <- scenarios %>% mutate(
  scenario=paste0(as.character(startbal/1000*-1),"k-",as.character(draw/1000),"k-",as.character(return*100),"%-", as.character(yrs),"y")
)

sched_all_scenario <- data.frame()
for(s in 1:nrow(scenarios)){
  scen <- scenarios$scenario[s]
  start_bal <- scenarios$startbal[s]
  sched_all <- data.frame()
  for(y in 1:scenarios$yrs[s]){
    remain <- fv(r=scenarios$return[s], n=1, pv=start_bal, pmt=scenarios$draw[s], type=0)
    sched <- data.frame('scenario'=scen,
                        'year'=y,
                        'balance'=remain)
    sched_all <- bind_rows(sched_all, sched)
    start_bal <- remain*-1
  }
  sched_all_scenario <- bind_rows(sched_all_scenario, sched_all)
}

```

```{r}
pbar <- sched_all_scenario %>% ggplot(aes(x=year, y=balance, fill=scenario))+geom_col(position = position_dodge())+
  scale_y_continuous(labels=comma)+
  labs(title=paste0("How much left after ", max(scenarios$yrs), " years in different scenarios"),
       y="", x="year from start")

ggplotly(pbar)
```

```{r}
pline <- sched_all_scenario %>% ggplot(aes(x=year, y=balance, color=scenario))+geom_line()+
  geom_hline(yintercept=0)+
  scale_y_continuous(labels=comma)+
  labs(title=paste0("How much left after ", yrs, " years in different scenarios"),
       y="", x="year from start")

ggplotly(pline)
```

## Simulations

### Setup and Run Simulations

```{r}

## number of simulations
nsims <- 100
## number of years to simulate
yrs <- 20
## parameters for starting, return, draw - exact numbers determined in loop
## min and max starting balance
start_min <- 600000
start_max <- 900000
## return rate 
r_mean <- 0.03
r_sd <- 0.07
r_max <- 0.20
## draw 
draw_min <- 50000
draw_max <- 75000

sim_all <- data.frame()

for(s in 1:nsims){
  sim <- s
  bal_start <- round(runif(n=1, min=start_min, max=start_max),0)
  bal_start_set <- bal_start
  bal_start_no_draw <- bal_start
  sim_sched <- data.frame()
  for(y in 1:yrs){
    rrate <- round(min(rnorm(n=1, mean=r_mean, sd=r_sd), r_max), 3)
    draw <- round(runif(n=1, min=draw_min, max=draw_max),0)
    return <- round(bal_start*rrate, 0)
    bal_remain <- ifelse(bal_start>0,bal_start+return-draw,bal_start-draw)
    bal_no_draw <- round(bal_start_no_draw*(1+rrate),0)
    sim_yr <- data.frame('sim'=sim,
                        'year'=y,
                        'start'=bal_start,
                        'rrate'=rrate,
                        'return'=return,
                        'draw'=draw,
                        'balance'=bal_remain,
                        'bal_no_draw'=bal_no_draw)
    sim_sched <- bind_rows(sim_sched, sim_yr)
    bal_start <- bal_remain
    bal_start_no_draw <- bal_no_draw
  }
  sim_all<- bind_rows(sim_all, sim_sched)
}
sim_all$sim <- as.factor(sim_all$sim)
```

* Number of simulations: `r nsims` 
* Number of years: `r yrs`
* Starting balance: `r start_min` to `r start_max` (uniform dist.)
* Rate of return: average `r r_mean` ave, `r r_sd` std dev (normal dist. with max `r r_max`)
* Annual draw: `r draw_min` to `r draw_max` (uniform dist.)

### Visualize Progress Over Time

```{r}
psim <- sim_all %>% ggplot(aes(x=year, y=balance, color=sim))+geom_line()+
  geom_hline(yintercept=0)+
  scale_y_continuous(labels=comma)+
  theme(legend.position = 'none')+
  labs(title=paste0("How much left after ", yrs, " years in different simulations"),
       y="", x="year from start")

ggplotly(psim)
```

### Summarize

#### Distribution of End Balances

```{r}
chart_title <- "Distribution of End Balances"
sim_all %>% ggplot(aes(x=balance))+geom_histogram()+
  labs(title=chart_title)
```

#### What Really Counts: under water or not?

```{r}
sim_all <- sim_all %>% filter(year==20) %>% mutate(
  pos=ifelse(balance>0,"money","no money")
)

sim_all %>% ggplot(aes(x=pos))+geom_bar()+
  labs("How likely to run out of money???")
```


```{r}
sim_neg <- sim_all %>% filter(balance<=0)
sim_neg_yr <- sim_neg %>% group_by(sim) %>% summarize(yr=min(year))

sim_neg_yr %>% ggplot(aes(x=yr))+geom_bar()+
  labs(title="Distribution of Years when Balance = $0")
```

For those simulations that go below $0 with end balance.

```{r}
chart_title <- "Distribution of Returns in Sims"
hist1 <- sim_all %>% ggplot(aes(x=rrate))+geom_histogram()+
  geom_vline(xintercept=mean(sim_all$rrate), linetype='dotted')+
  scale_y_continuous(expand=expansion(add=c(0,1)))+
  scale_x_continuous(labels=percent)+
  labs(title=chart_title)

mrate_all <- mean(sim_all$rrate)
sdrate_all <- sd(sim_all$rrate)

## calc returns when draw downs excluded
sim_rr <- sim_all %>% group_by(sim) %>% summarize(start_bal=first(start),
                                                  end_bal=last(bal_no_draw),
                                                  yrs=max(year)) %>%
  mutate(ttl_return=end_bal/start_bal-1,
         ttl_return_ave=ttl_return/yrs)

chart_title <- "Distribution of Real Returns in Sims (no draw)"
hist2 <- sim_rr %>% ggplot(aes(x=ttl_return_ave))+geom_histogram()+
  geom_vline(xintercept=mean(sim_rr$ttl_return_ave), linetype='dotted')+
  scale_y_continuous(expand=expansion(add=c(0,1)))+
  scale_x_continuous(labels=percent)+
  labs(title=chart_title)

grid.arrange(hist1, hist2, nrow=1)
```

* Mean rate of return (straight ave): `r mrate_all`
* Std dev rate of return (straight sd): `r sdrate_all`

* Mean total rate of return, no draws: `r mean(sim_rr$ttl_return_ave)`
* Std dev rate of return, no draws: `r sd(sim_rr$ttl_return_ave)`


```{r}
chart_title <- "Returns over time"
sim_all %>% ggplot(aes(x=as.factor(year), y=return))+geom_boxplot()+
  geom_hline(yintercept = 0)+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title)
```

```{r}
chart_title <- "Draws over time"
sim_all %>% ggplot(aes(x=as.factor(year), y=draw))+geom_boxplot()+
  geom_hline(yintercept=mean(sim_all$draw), linetype='dashed')+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title, x='year from start')
```

### Evaluate Positive Outcomes

#### Identify Positive Outcomes

```{r}
## identify sims where balance > 0 at year 20
sim_pos <- sim_all[sim_all$year==20 & sim_all$balance>0,]
sim_pos_rate <- nrow(sim_pos)/nsims
## filter full list for positive sims only
sim_pos_filter <- sim_pos %>% select(sim)
sim_pos_all <- left_join(sim_pos_filter, sim_all, by='sim')

## summarize individual sims
sim_pos_ind_smry <- sim_pos_all %>% group_by(sim) %>% summarize(
  start_bal=first(start),
  ave_rate=round(mean(rrate),3),
  ave_draw=round(mean(draw),0),
  end_bal=last(balance)
)

chart_title <- paste0("Ending Balance after ", yrs," yrs for positive sims")
sim_pos %>% ggplot(aes(x=reorder(sim, -balance), y=balance))+geom_col(position=position_dodge())+
  scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.05)))+
  labs(title=chart_title, x='sim')
```

#### Overall Sim Summary

```{r}
sim_pos_smry <- sim_pos_ind_smry %>% summarize(
  ave_start=mean(start_bal),
  ave_rate=mean(ave_rate),
  ave_draw=mean(ave_draw),
  end_bal=mean(end_bal)
)
sim_pos_smry
```

#### Individual Pos Sim Summary

```{r}
sim_pos_ind_smry 
```

#### Detailed Table

```{r}
## DT to add table of all rows for each positive sim for inspection
datatable(sim_pos_all)
```

#### What is distribution in returns within sims?

```{r}
chart_title <- 'Distributions in Annual Returns across positive sims'
sim_pos_all %>% ggplot(aes(x=sim, y=rrate))+geom_boxplot()+
  geom_hline(yintercept = 0)+
  labs(title=chart_title)
```

#### Is there correlation between starting balance & returns?

Do sims in the positive set have both high starting AND high returns?

```{r}
sim_pos_ind_smry %>% ggplot(aes(x=start_bal, y=ave_rate))+geom_point()
```
