---
title: "Scenario explore"
author: "JY"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(lubridate)
library(scales)
library(plotly)

library(xts)
library(zoo)
library(quantmod)
library(Quandl)
library(dygraphs)
library(PerformanceAnalytics)

## theme set for ggplot
theme_set(theme_light())
```

## Objective

Explore some future scenarios based on past results in financial markets. This will help get a sense of some realistic future possibilites for fictional portfolio.

## Step 1. Get Historical Returns

* S&P 500: monthly returns

(this work already done in 'SP500.Rmd')

```{r}
## tip: set start date mth before you want data so that you will have mthly returns for full period
## (since there is no return calc for the first mth, by def)
sp500 <- getSymbols(Symbols="^GSPC", from="1969-12-01", auto.assign=FALSE)
head(sp500)
```

```{r}
sp500_mth <- to.monthly(sp500)

sp500_mth_ret <- Return.calculate(sp500_mth$sp500.Close)

chart_Series(sp500_mth_ret)
```

* Vanguard all-stocks: monthly returns

```{r}

```

## Step 2. Distribution of Returns

* Identify distribution of historical monthly returns
* Assuming approx normal distribution, identify:
    + mean
    + std dev

```{r}
## drop first mth, which is NA (by def)
sp500_mth_ret <- sp500_mth_ret[-1,]

sp500_mth_ret_ave <- mean.geometric(sp500_mth_ret)

sp500_mth_ret_stdev <- StdDev(sp500_mth_ret) ## StdDev from PerformanceAnalytics
sp500_mth_min <- min(sp500_mth_ret, na.rm=TRUE)
sp500_mth_max <- max(sp500_mth_ret, na.rm=TRUE)

## annualized return over whole period
sp500_ttl_ann_ret <- Return.annualized(sp500_mth_ret, scale=12)

```


## Step 3. Monte Carlo Simulation

* Generate a monthly return for each month by random selection from the distribution identified above
* Calculate future value based on this selection of returns
* Repeat entire process multiple times to understand the expected range of possibilities

alternative:

* could look at rolling 5-yr, 7-yr, 10-yr performance (by month)
* use that distribution to create range of values for performance for those periods (rather than individual monthly returns)

```{r}
amount <- 400000
nmth <- 60
nsims <- 40

returns_all <- data.frame()

for(s in 1:nsims){
  returns <- data.frame(mth=NA, ret=NA)
  for(r in 1:nmth){
    returns[r,"mth"] <- r
    returns[r,"ret"] <- rnorm(n=1, mean=sp500_mth_ret_ave[1], sd=sp500_mth_ret_stdev[1]) 
  }
  returns <- returns %>% mutate(yr=ceiling(mth/12)
  )
  returns <- returns %>% mutate(
    sim=as.factor(s)
  )
  returns$amt <- NA
  for(rr in 1:nrow(returns)){
    returns[rr,'amt'] <- ifelse(rr==1,amount*(1+returns[rr,'ret']), returns[rr-1,'amt']*(1+returns[rr,'ret']))
  }
  returns_all <- bind_rows(returns_all, returns)
}

```

Visualize balance over time

```{r}
chart_title <- paste0('Monthly Returns for ',nsims,' simulations of ',nmth, ' months')
returns_all %>% ggplot(aes(x=mth, y=amt, color=sim))+geom_line()+
  scale_y_continuous(labels=comma)+
  theme(legend.position = 'none')+
  labs(title=chart_title)
```

```{r}
## ave final amount
returns_amt <- returns_all %>% filter(mth==nmth) %>% select(mth, sim, amt)

chart_title <- paste0("Distribution of End Amt for ",nsims, " of ",nmth, " months") 
plot_dist <- returns_amt %>% ggplot(aes(x=amt))+geom_histogram()+
  scale_x_continuous(labels=comma)+
  geom_vline(xintercept=mean(returns_amt$amt), color='blue')+
  geom_vline(xintercept=median(returns_amt$amt), color='green')+
  geom_vline(xintercept=quantile(returns_amt$amt, 0.25))+
  geom_vline(xintercept=quantile(returns_amt$amt, 0.75))+
  labs(title=chart_title)
ggplotly(plot_dist)
```

```{r}
median(returns_amt$amt)
mean(returns_amt$amt)
quantile(returns_amt$amt, 0.25)
quantile(returns_amt$amt, 0.75)

```

