---
title: "Stock Analysis with PerformanceAnalytics"
author: "`r Sys.getenv('USER')`"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=3.5,
                      fig.width=6)

library(tidyverse)
library(lubridate)
library(here)
library(plotly)
library(scales)
library(gridExtra)
library(DT)

library(PerformanceAnalytics)
library(quantmod)
library(xts)
library(dygraphs)

```

## Intro

Examples of return and portfolio analysis that can be done with the PerformanceAnalytics package. A basic starting point, **far from comprehensive**.

* [Performance Analytics pkg manual](https://cran.r-project.org/web/packages/PerformanceAnalytics/PerformanceAnalytics.pdf) - very thorough documentation but the challenge is it is arranged alphabetically by function (as typical of R vignettes). Works best if you already know what you are looking for, rather than a linear learning tool. There is an intro section that provides some lay-of-the-land for returns-based-analysis (RBA).
* [Blog post: Performance Analytics: An Indispensable Quant Tool for Any Investor](https://franklinparker.com/2019/02/16/performanceanalytics-an-indespensible-quant-tool-for-any-investor/) - brief but helpful tips on getting started.

## Get Price Data

Select 3 ETFs to compare for the purpose of selection.

```{r}
etf_symbols <- c("^GSPC","VTI","VIG","ZLB.TO")
## set empty object
etf_all <- NULL
## loop through each stock, binding cols as you go
for(stocks in etf_symbols){
  etf_all <- cbind(etf_all,
                      getSymbols(Symbols=stocks, 
                                 from='2018-01-01', periodicity='daily', 
                                 auto.assign = FALSE))
}
## remove any NA rows


head(etf_all)

```

### Initial viz

* Prices: not that useful because scale comes into play, and is irrelevant

```{r}
dygraph(Ad(etf_all))
```

## Normalized price ranges for comparison

To enable better comparison of prices, we can normalize them with min-max-normalization to set them all between 0 and 1. 

From chatGPT:
"Normalizing between 0 and 1 (also called min-max normalization) scales the variables to have a minimum value of 0 and a maximum value of 1. This method is useful when you want to preserve the original range of the data and the relationship between the values."

```{r}
## first remove any NA values - they will mess up calculations
etf_all <- na.omit(etf_all)
## use apply to calc min-max normalization
etf_all_mmn <- xts(apply(etf_all, 2, function(x) (x-min(x))/(max(x)-min(x))), 
                   order.by=index(etf_all))
```

```{r}
dygraph(Ad(etf_all_mmn))
```

More useful view where we can see the tight connection in prices - and instances where the connection is not so tight.

Still, to get to the heart of the matter, we need to look at **returns**.

## Get Returns

'cause in the end: **returns are what matter** ;)

**Performance Analytics pkg** focuses on returns analysis. From the manual:

*"Performance measurement starts with returns... Most of the recent work
in performance analysis is focused on returns rather than prices and sometimes called
"returns-based analysis" or RBA. This “price per unit of investment” standardization is important
for two reasons - first, it helps the decision maker to compare opportunities, and second, it has some
useful statistical qualities."*

### Convert price data to returns

Using monthly returns in this case.

```{r}
## potentially simpler way to calc monthly returns
## calc only on adjusted prices - every 6th col
etf_mth_ret <- NULL
for(i in 1:length(etf_symbols)){
  cadj <- i*6
  etf_mr <- monthlyReturn(etf_all[,cadj])
  colnames(etf_mr) <- etf_symbols[i]
  etf_mth_ret <- cbind(etf_mth_ret, etf_mr)
}
```

## Correlations

```{r}
chart.Correlation(etf_mth_ret)
```

From [Performance Analytics: An Indispensable Quant Tool for Any Investor](https://franklinparker.com/2019/02/16/performanceanalytics-an-indespensible-quant-tool-for-any-investor/)

"The diagonal gives you the return distribution of each ETF. I like to see how the returns are distributed—are they concentrated in the center? are there lots of outliers? are they heavily skewed? is something just weird? Indeed, this plot recently helped me immediately throw out a position under consideration because the distribution was barbelled (but more heavy left-tailed)."

## Distribution of Returns

Focus on individual distribution of returns:

* line is normal curve
* shows 95% value at risk level (incl. modified)

```{r}
## function for producing individual dist
## - provide symbol to select (by number in list), list of symbols (from start), dataset of returns as data frame
fn_ind_rtn_dist <- function(sel, symbs, data){
  chart.Histogram(data[,sel],
                methods=c("add.normal","add.risk"),
                colorset = c('steelblue','','navyblue'),
                main=symbs[sel])
}
```

```{r}
nsymb <- length(etf_symbols)
mfr <- ifelse(nsymb<4,1,round(nsymb/2))
mfc <- min(nsymb,3)
par(mfrow=c(mfr,mfc))
for(c in 1:length(etf_symbols)){
 fn_ind_rtn_dist(c, etf_symbols, etf_mth_ret)  
}

```

## Drawdown comparisons

Shows how resilient investment is during negative return situations.

* downward movement relative to a high.

```{r}
#sel <- 1
#chart.Drawdown(etf_mth_ret_cl[,sel], geometric=TRUE, colorset="firebrick4", main=etf_symbols[sel])
## plot.engine options: ggplot2, plotly, dygraph, googlevis, default
chart.Drawdown(etf_mth_ret, geometric=TRUE, legend.loc='bottomleft', plot.engine="dygraph")

```

## Upside/Downside Capture

Shows the relative percentage of upside and downside captured by an asset relative to a benchmark.

```{r}
chart.CaptureRatios(Ra=etf_mth_ret[,2:4], Rb=etf_mth_ret[,"X.GSPC"], colorset="dodgerblue4",
    main="Capture Ratio")
```

Couple things to note:

* the benchmark will be situated where the x and y axes cross at 1. This is because it captures 100% of the upside and downside of itself.
* other assets are then evaluated by how far and what direction they sit on both x and y axis from the benchmark.
* downside that is lower than 1 means the asset tends to not to react with as much losses on the downside as the benchmark.
* upside greater than 1 means the asset tends to see greater gains on the upside than the benchmark.

## Conclusion

Lots to work with in the PerformanceAnalytics package - these examples are not even scratching the surface. Likely have further editions to come. 