---
title: "Stock-performance-basic"
author: "John Yuill"
date: "`r Sys.Date() `"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      toc_collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(lubridate)
library(scales)

library(xts)
library(zoo)
library(quantmod)
library(Quandl)
library(dygraphs)

## theme set for ggplot
theme_set(theme_light())
```

## Get price data

Getting daily price data for any stock is as easy as using quantmod package and getSymbols function.

* fetches data as xts time series object
* note src and updated in str output 
* auto.assign:
    - TRUE: creates xts object in environment with name of symbol
    - FALSE: prints results OR can be used to save results to object of any chosen name

```{r}
getSymbols("CSIQ", auto.assign=TRUE) ## auto.assign=TRUE saves object
csiq_prices <- getSymbols("MSFT", auto.assign=FALSE) ##auto.assign=FALSE to save to specified object
## xts object
str(CSIQ)
```
Entire history of the daily stock price (at least going back to 2007) is now available in an xts object named after the stock symbol. 

* default source is yahoo finance (src='yahoo')
* default start date is '2007-01-01' (from='2007-01-01')

Provides daily Open, High, Low, Close, Volume, Adjusted (aka OHLCVA):

```{r}
## show first rows 
head(CSIQ)
## show last rows
tail(CSIQ)
```

"Adjusted" = prices adjusted for stock splits and dividends (possibly other developments? major new issues? buy-backs?)

### Incl. TSX

* Includes stocks on TSX
* Optionally import into specified object.
* Specify src from variety of available data sources

```{r}
## specify object with auto.assign=FALSE
ogi <- getSymbols(Symbols='OGI.TO', auto.assign=FALSE, src='yahoo')
```

### Filter for time period

* When importing: specific period (start date and end date)

```{r}
getSymbols("CSIQ", from="2018-01-01", to=Sys.Date(), auto.assign = TRUE)
CSIQ_2018 <- CSIQ

head(CSIQ_2018)

## reset full CSIQ for later
getSymbols("CSIQ", auto.assign = TRUE)

```

* filter after import

```{r}
CSIQ_2020 <- CSIQ['2020-01-01/'] ## from specified date fwd
CSIQ_90d <- tail(CSIQ, n=90) ## last n 

## selected period
CSIQ_Aug <- CSIQ['2020-08/2020-08-31'] ## Aug 1-31
```

* filter with first and last functions in xts

```{r}
## data in last 3 months (calendar mths, data so far)
CSIQ_3_mth <- last(CSIQ, "3 months" )
## combine first & last to get: first week from last (mont recent) year
## - includes the days of week 1 in the year
CSIQ_first_week <- first(last(CSIQ, "1 year"), "1 week")
```

More Time Series Manipulation examples with xts below in separate section.

### Get multiple stocks 

Each imported as separate object

```{r}
getSymbols(c("NEP","SPWR"), auto.assign = TRUE)
## OR
stocklist <- c("NEP","SPWR","FSLR")
getSymbols(stocklist, aauto.assign = TRUE)

```

### Merge multiple stocks

* use merge function to join cols
* max 2 data sets at a time, unless use do.call

```{r}
## TWO AT A TIME
## first two
renewables <- merge(NEP, SPWR, join='outer', fill=NA)
head(renewables)
## add another one
renewables <- merge(renewables, FSLR, join='outer', fill=NA)
head(renewables)

## MULTIPLE (2+) AT A TIME
## set list of symbols
stockrenew <- c("BEP","TERP","VSLR")
## set up a new environment to hold the imported data
data_env <- new.env()
## get symbols
getSymbols(Symbols=stockrenew, env=data_env, auto.assign=TRUE)
## create list of objects within the environment
data_list <- lapply(data_env, Cl) ## apply function: Cl=close; OHLCV (not sure how to get all)
renewables2 <- do.call(merge, data_list)
head(renewables2)
```

### Multiple - most efficient

Most efficient code for getting multiple stocks at once.

* use cbind to combine cols in separaet object as you go 

```{r}
## specify stock symbols
stockrenew3 <- c("ENPH","SEDG","RUN","JKS")
## set empty object
renewables3 <- NULL
## loop through each stock, binding cols as you go
for(stocks in stockrenew3){
  renewables3 <- cbind(renewables3,
                      getSymbols(Symbols=stocks, 
                                 from='2018-01-01', periodicity='daily', 
                                 auto.assign = FALSE))
}
head(renewables3)

```

* CLOSE only: specify [,4] at end of getSymbols

```{r}
## get CLOSE only: same as above but SPECIFY [,4] at end
renewables3c <- NULL
for(stocks in stockrenew3){
  renewables3c <- cbind(renewables3c,
                      getSymbols(Symbols=stocks, 
                                 from='2018-01-01', periodicity='daily', 
                                 auto.assign = FALSE)[,4])
}
head(renewables3c)
```

### Data frame conversion

```{r}
## convert xts to data frame - will work with dygraphs
csiq_df <- data.frame(csiq_prices)
## row names will be dates - move to date col
csiq_df$date <- ymd(rownames(csiq_df))
## [opt] set rownames to number as in typical data frame (won't work with dygraphs)
rownames(csiq_df) <- 1:nrow(csiq_df)
## [opt] move date col to first left
csiq_df <- csiq_df %>% select(date,1:(ncol(csiq_df)-1))
```


## Visualize 

quantmod has its own 'chartSeries' / 'chart_Series', Dygraphs pkg enables interactive charts designed to work with quantmod right out of the box.

Several options:

* **chartSeries**: basic charting from Quantmod pkg
    - works with xts
    - understands .Open, .High, .Low, .Close
    - basic version has vol
    - easy to add technical layers
* **chart_Series**: experimental version of basic charting from Quantmod pkg, seems like similar features, different default formatting
* **dygraphs**: more interactivity, works with quantmod, **xts**
    - works with data frame as long as rownames are dates 
    - doesn't have the technical analysis features (afaik); limited tweaking of format - mostly automatic
* **ggplot**: if data converted from xts to data frame (example above) 

### chartSeries: basic line charts (non-interactive)

#### Candlestick

* chartSeries, chart_Series: single stock, OHLC candlestick by default

```{r CHART_SERIES_ONE}
## basic chart_Series from quantmod
## - two versions chartSeries (orig) and chart_Series (new experimental)
chartSeries(CSIQ_90d, name="CSIQ - OHLC & vol")
chart_Series(CSIQ_90d, name="CSIQ - OHLC") 
## same:
#chart_Series(OHLC(CSIQ_90d))
```

* Open is either bottom of green bar OR top of red bar
* High is highest point (either line or top of bar)
* Low is lowest point (either line or bottom of bar)
* Close is either top of green bar OR bottom of red bar

#### Technical features

* SMA: simple moving average
* MACD: moving average convergence divergence signals
* BBands: Bollinger bands
* other features are available in quantmod (not shown)

```{r}
chartSeries(CSIQ_90d, name="CSIQ - OHLC & Technical Indicators", TA=c(addVo(),addBBands(), addMACD()))
```

* add more technical with chart_Series - layer in one chart at a time

```{r}
chart_Series(OHLC(CSIQ), subset='2019-12/2020-07')
add_SMA(n=30, col='blue')
add_SMA(n=7, col='green')
add_MACD() 
add_BBands()
addMomentum()
```

* same idea all-in-one (leaving out momentum - doesn't work with chart_Series)

```{r}
## semi-colon separated TA functions
chart_Series(OHLC(CSIQ), subset='2019-12/2020-07', main="Full Tech Indicators",
             TA="add_SMA(n=30, col='blue');
             add_SMA(n=7, col='green');
             add_MACD();
             add_BBands()")
```


#### Specific cols

* can isolate specific columns

```{r}
## two variations
chartSeries(Cl(CSIQ_90d), name="CSIQ - Closing Prices")
## new charting
chart_Series(Cl(CSIQ_90d), name="CSIQ - Closing Prices")
```

#### Multiple stocks

* multiple stocks, one chart (but how to identify each stock?)

```{r}
chart_Series(Cl(renewables3c))
chartSeries(renewables, theme=chartTheme("white"))
```

### Dygraph line/candle (interactive)

* dygraph is designed for time series data - xts OR data frame, with first col as date
* works with quantmod but doesn't incorporate separate vol or technical features
* for OHLC need dyCandlestick additional command
* (Info on inputs for dygraphs)[https://r-graph-gallery.com/316-possible-inputs-for-the-dygraphs-library.html]

#### Lines / Candlestick

```{r DYGRAPH_ONE}
## use dygraphs pkg
dygraph(OHLC(CSIQ_90d))
cndl <- dygraph(OHLC(CSIQ_90d))
dyCandlestick(cndl)

```

#### Single col

* closing price only

```{r}
# chart_Series(Cl(CSIQ)) #chart_Series seems to sometimes cause probs with pandoc(?)
dygraph(Cl(CSIQ))

```

#### Annotations!

Easily add annotations

```{r}
## shorter date range for readability / relevance
CSIQ_covid <- CSIQ_2020['/2021-05-30']
## find date of peak close for annotation
#max(CSIQ_covid$CSIQ.Close)
peak_date <- index(CSIQ_covid[Cl(CSIQ_covid)==max(Cl(CSIQ_covid))])

## multi-step process
## setup chart parameters
#graph <- dygraph(OHLC(CSIQ_covid), main="Cdn Solar")
## compile/add event annotations
#graph <- dyEvent(graph, c("2020-02-21",as.character(peak_date)),
#                          c("Start of covid crash","Post-covid peak"),
#                 labelLoc="top")
## print candlestick
#dyCandlestick(graph)

## pipe together chart info and event annotations individually for more control
dygraph(OHLC(CSIQ_covid), main="Cdn Solar with annotation") %>%
  dyEvent("2020-02-21","Start of covid crash", labelLoc="top") %>%
  dyEvent(as.character(peak_date),"Post-covid peak", labelLoc="bottom") %>%
  dyCandlestick()
```

Show only adjusted close, filtered for 2020 - early 2021

```{r}
CSIQ_2020 <- CSIQ['2020-01-01/2021-01-31']
graph <- dygraph(Ad(CSIQ_2020), main="Cdn Solar")
graph <- dyEvent(graph, "2020-02-21","Start of covid crash",
                 labelLoc="top")
graph <- dyEvent(graph, "2021-01-21","Peak after covid crash",
                 labelLoc="bottom")
graph
```

* just under 1 yr from crash to peak

## Adjusted prices

Adjusted price is supposed to adjust for stock splits, while other prices are actual on the trading day.

However...Tesla (TSLA) had a 5-for-1 stock split on Aug 3 dropping price from ~$2,000 to ~$400. Prices collected as of Sep 5, 2020 show post-split throughout.

```{r}
getSymbols("TSLA", auto.assign = TRUE)

## filter for Aug, pre-split
TSLA_presplit <- TSLA['2020-08/2020-08-30']
tail(TSLA_presplit)

```

```{r}
#dygraph(OHLC(TSLA))
## set width and height in pixels
dygraph(Ad(TSLA), width=400, height=300)
dygraph(TSLA[,c(4,6)], width=400, height=300)
```

## Get RETURNS

* Prices are great, but **what REALLY counts is RETURNS!**
* Quantmod has functions for different returns by periodicity (daily, weekly, monthly, quarterly, annual)
* **periodReturn** is the underlying function
* can use **allReturns** to calculate all return periods at once, depending on date range.
* **PerformanceAnalytics** pkg is the definitive R pkg for returns and portfolio optimization - so this is quick coverage and will dive more into PerformanceAnalytics in another file.

```{r}
CSIQ_90d_return <- dailyReturn(CSIQ_90d)
head(CSIQ_90d_return)
dygraph(CSIQ_90d_return, width=600, height=250)
```

#### All Returns data

* one command to calculate returns for all periodicities - assuming date range is wide enough.

```{r}
CSIQ_returns <- allReturns(CSIQ)
head(CSIQ_returns)
```

#### Annual returns

* zoom in on annual returns.

```{r}
## subset for annual return data and chart
CSIQ_returns[!is.na(CSIQ_returns$yearly),5] %>%
dygraph(main="CSIQ Annual Returns")
head(CSIQ_returns[!is.na(CSIQ_returns$yearly),5])
```

* would be nicer to have bar chart
* how about quick **ggplot after conversion to data frame**?

```{r, fig.height=3, fig.width=6}
csiq_ret_df <- data.frame(CSIQ_returns[!is.na(CSIQ_returns$yearly),5])
csiq_ret_df$date <- ymd(rownames(csiq_ret_df))
ggplot(csiq_ret_df, aes(x=as.factor(date), y=yearly))+geom_col()+
  scale_y_continuous(labels=percent_format())+
  labs(title="Yearly Returns", x="", y="annual % returns")+
  theme(axis.text.x = element_text(angle=45, hjust=1))
## or simplified x-axis
csiq_ret_df$year <- year(csiq_ret_df$date)
csiq_ret_df %>% ggplot(
  aes(x=year, y=yearly))+geom_col()+
  scale_y_continuous(labels=percent_format())+
  labs(title="Yearly Returns", x="", y="annual % returns")

```


### Stats on returns

```{r}

summary(CSIQ_90d_return)
sum(CSIQ_90d_return)

ggplot(CSIQ_90d_return, aes(x=CSIQ_90d_return))+geom_histogram()+
  geom_vline(xintercept = median(CSIQ_90d_return))+
  geom_vline(xintercept = quantile(CSIQ_90d_return, 0.2))+
  geom_vline(xintercept = quantile(CSIQ_90d_return, 0.8))+
  scale_y_continuous(expand=c(0,0))+
  scale_x_continuous(labels=percent)+
  labs(title="Daily returns over the period", 
       subtitle='Lines: 20th percentile, Median, 80th percentile',
       x='Daily % chg')

```

## Time Series Manipulation

Leverage xts to work with time series objects (probably warrants its own file)

### End points by period

* day, week, month, qtr, yr.
* returns the locations of each last item in the specified type of period, aligned on calendar.
* will include the last item available in the last period (last data point).
* returns a vector starting with 0 and ending with the extent (last row) of your data.
* can also use k to skip: on='years', k=4 returns location of every 4th yr in data (plus last)

```{r}
## get index number of last day of each year
index_yr <- endpoints(CSIQ, on="years")
## apply to time series to extract selected end points
CSIQ_yr_end <- CSIQ[index_yr]
head(CSIQ_yr_end[,c(2,3,5)]) ## selected cols for space

```

```{r}
index_week_end <- endpoints(CSIQ_Aug, on="weeks")
CSIQ_Aug[index_week_end, 1:4] ## open, high, low, close selected
```

### Summarization

```{r}
##
period.apply(CSIQ_Aug, INDEX=index_week_end, FUN=mean)

```

### Shortcut summarization

```{r}
## apply.weekly, apply.monthly, apply.quarterly, apply.yearly to summarize
## FUN=<whatever summarization function for the period)
apply.weekly(CSIQ_Aug, FUN=mean)
```


### Split into periods

```{r}
## split data into weeks - chunk the data into 'week' components, without summarizing
CSIQ_Aug_weeks <- split(CSIQ_Aug, f='weeks')
## second week: high, low, close selected
CSIQ_Aug_weeks[[2]][,2:4]
```

## Aggregate time periods

* Aggregate data by period
* Can take a single variable and calculate OHLC statistics by period

```{r}
CSIQ_yearly <- to.period(CSIQ[,6], period='years', name='CSIQ_adj', OHLC=TRUE)
CSIQ_yearly
```

### Shortcut aggregation

```{r}
CSIQ_2020_mth_agg <- to.monthly(CSIQ_2020[,6], name='CSIQ_adj')
CSIQ_2020_mth_agg
```


### Merging data with difference frequencies

```{r}
## get data -> DFF is daily Fed rate, FEDFUNDS is monthly
getSymbols(Symbols=c("FEDFUNDS","DFF"), src="FRED" )

head(DFF) ## daily data 
head(FEDFUNDS) ## mthly data
```

```{r}
## aggregate DFF to monthly by average
DFF_monthly <- apply.monthly(DFF, FUN=mean) ## FUN=aggregation calc
head(DFF_monthly) ## date as of end of month

## convert to monthly index
index(DFF_monthly) <- as.yearmon(index(DFF_monthly))

## mow can merge with data sets that have first of month
merged_funds <- merge(FEDFUNDS, DFF_monthly)
head(merged_funds)
```

## Quandl for Importing

Claims 'the world's most powerful data lives on [Quandl](https://www.quandl.com/). See also: [Quandl API documentation](https://docs.quandl.com/docs/getting-started).

* Economic data and similar
* Specify Source/Symbol
* Several 'types' available: deflaut is data.frame
* Does not appear that you can access stock prices without account/payment :(

```{r}
## as of Jan 14, 2023 getting error msg that I need to get API key
#us_gdp <- Quandl(code='FRED/GDP')

#ggplot(us_gdp, aes(x=Date, y=Value))+geom_line()
```

## PerformanceAnalytics [next]

PerformanceAnalytics pkg is considered best for analysis of returns and portfolio decisions. Will dive into it in separate file.