---
title: "Stock-performance-check"
author: "John Yuill"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      toc_collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(lubridate)
library(scales)

library(xts)
library(zoo)
library(quantmod)
library(Quandl)
library(dygraphs)

## theme set for ggplot
theme_set(theme_light())
```

## Get price data

Getting daily price data for any stock is as easy as using quantmod package and getSymbols function.

* fetches data as xts time series object
* note src and updated in str output 

```{r}
#symb <- "T" ## AT&T
symb <- "CVS" ## CVS pharmacy
stock <- getSymbols(symb, auto.assign=FALSE) ## auto.assign=TRUE saves object as symbol

## clean up colnames by removing symbol for versatility
colnames(stock) <- str_replace(colnames(stock),str_glue(symb,"."), "")
```


### Filter for time period

```{r}
stock_2020 <- stock['2020-01-01/'] ## from specified date fwd
#stock_52w <- tail(stock, n=365) ## last n 
stock_52w <- last(stock, "52 weeks")

## selected period
stock_90d <- last(stock, "3 months") ## goes back to start of 3rd mth previous
```

### Summary stats

```{r}

h52w <- max(stock_52w$High)
l52w <- min(stock_52w$Low)
l52wa <- min(stock_52w$Adjusted)
cl <- last(stock_52w$Close, "1 day")

## create data frame version of xts for flexibility
ds_stock52w <- as.data.frame(stock_52w)
## get dates from index as field
ds_stock52w$date <- rownames(ds_stock52w)

ds_stock52w %>% ggplot(aes(x=Adjusted))+geom_histogram()+
  geom_vline(xintercept=cl)+
  ggtitle(paste0(symb," price distribution"))

```



## Visualize 

Dygraphs pkg enables interactive charts designed to work with quantmod right out of the boox.

```{r}
dygraph(Ad(stock_52w))
```


### Basic line chart

* with volume

```{r}
barChart(stock_52w)
```

## Add Technical features

* SMA: simple moving average
* MACD: moving average convergence divergence signals
* BBands: Bollinger bands
* other features are available in quantmod (not shown)

```{r}
chart_Series(Cl(stock_52w))
add_SMA(n=90, col='blue')
add_SMA(n=180, col='green')
add_MACD() 
add_BBands()
addMomentum()
```

## Get Returns

```{r}
stock_52w_return <- dailyReturn(stock_52w)
dygraph(stock_52w_return)
```

### Stats on returns

```{r}

summary(stock_52w_return)
sum(stock_52w_return)

ggplot(stock_52w_return, aes(x=stock_52w_return))+geom_histogram()+
  geom_vline(xintercept = median(stock_52w_return))+
  geom_vline(xintercept = quantile(stock_52w_return, 0.2))+
  geom_vline(xintercept = quantile(stock_52w_return, 0.8))+
  scale_y_continuous(expand=c(0,0))+
  scale_x_continuous(labels=percent)+
  labs(title="Daily returns over the period", 
       subtitle='Lines: 20th percentile, Median, 80th percentile',
       x='Daily % chg')

```


