---
title: "Stock-performance-basic"
author: "John Yuill"
date: "August 29, 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      toc_collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(lubridate)
library(scales)

library(xts)
library(zoo)
library(quantmod)
library(dygraphs)

## theme set for ggplot
theme_set(theme_light())
```

## Get price data

Getting daily price data for any stock is as easy as using quantmod package and getSymbols function.

```{r}
getSymbols("CSIQ", auto.assign=TRUE) ## auto.assign=TRUE saves object; FALSE prints out data

```

Entire history of the daily stock price (at least going back to 2007) is now available in an xts object named after the stock symbol. 

Provides daily Open, High, Low, Close, Volume, Adjusted (aka OHLCVA):

```{r}
## show first rows 
head(CSIQ)
## show last rows
tail(CSIQ)
```

"Adjusted" = prices adjusted for stock splits and dividends (possibly other developments? major new issues? buy-backs?)

### Filter for time period

* When importing: specific period (start date and end date)

```{r}
getSymbols("CSIQ", from="2018-01-01", to=Sys.Date(), auto.assign = TRUE)
CSIQ_2018 <- CSIQ

head(CSIQ_2018)

## reset full CSIQ for later
getSymbols("CSIQ", auto.assign = TRUE)
```

* filter after import

```{r}
CSIQ_2020 <- CSIQ['2020-01-01/'] ## from specified date fwd
CSIQ_90d <- tail(CSIQ, n=90) ## last n 

## selected period
CSIQ_Aug <- CSIQ['2020-08/2020-08-31'] ## Aug 1-31
```

* filter with first and last functions in xts

```{r}
## data in last 3 months (calendar mths, data so far)
CSIQ_3_mth <- last(CSIQ, "3 months" )
## combine first & last to get: first week from last (mont recent) year
## - includes the days of week 1 in the year
CSIQ_first_week <- first(last(CSIQ, "1 year"), "1 week")
```

More Time Series Manipulation examples with xts below in separate section.

## Visualize 

Dygraphs pkg enables interactive charts designed to work with quantmod right out of the boox.

### Basic line chart

* chart_Series 

```{r CHART_SERIES_ONE}
## basic chart_Series from quantmod
chart_Series(OHLC(CSIQ))

```

* dygraph

```{r DYGRAPH_ONE}
## use dygraphs pkg
dygraph(OHLC(CSIQ))

```

* with volume

```{r}
barChart(CSIQ)
```

* closing price only

```{r}
# chart_Series(Cl(CSIQ)) #chart_Series seems to sometimes cause probs with pandoc(?)
dygraph(Cl(CSIQ))

```

### Get multiple stocks 

Each imported as separate object

```{r}
getSymbols(c("NEP","SPWR"), auto.assign = TRUE)
## OR
stocklist <- c("NEP","SPWR","FSLR")
getSymbols(stocklist, aauto.assign = TRUE)
```


Easily add annotations

```{r}
graph <- dygraph(OHLC(CSIQ), main="Cdn Solar")
graph <- dyEvent(graph, "2020-02-21","Start of covid crash",
                 labelLoc="top")
graph

```

Show only adjusted close, filtered for 2020

```{r}
CSIQ_2020 <- CSIQ['2020-01-01/']
graph <- dygraph(Ad(CSIQ_2020), main="Cdn Solar")
graph <- dyEvent(graph, "2020-02-21","Start of covid crash",
                 labelLoc="top")
graph

```

## Adjusted price

Adjusted price is supposed to adjust for stock splits, while other prices are actual on the trading day.

However...Tesla (TSLA) had a 5-for-1 stock split on Aug 3 dropping price from ~$2,000 to ~$400. Prices collected as of Sep 5, 2020 show post-split throughout.


```{r}
getSymbols("TSLA", auto.assign = TRUE)

## filter for Aug, pre-split
TSLA_presplit <- TSLA['2020-08/2020-08-30']
tail(TSLA_presplit)

```

```{r}
#chart_Series(OHLC(TSLA))
dygraph(OHLC(TSLA))
```


## Candlestick

With selected period - in this case, based on recent (suitable up to ~30 days at a time).

```{r}
CSIQ_recent <- tail(CSIQ, n=30)
graph <- dygraph(OHLC(CSIQ_recent))
dyCandlestick(graph)

```

* Open is either bottom of green bar OR top of red bar
* High is highest point (either line or top of bar)
* Low is lowest point (either line or bottom of bar)
* Close is either top of green bar OR bottom of red bar

## Add Technical features

* SMA: simple moving average
* MACD: moving average convergence divergence signals
* BBands: Bollinger bands
* other features are available in quantmod (not shown)

```{r}
chart_Series(OHLC(CSIQ), subset='2019-12/2020-07')
add_SMA(n=30, col='blue')
add_SMA(n=7, col='green')
add_MACD() 
add_BBands()
addMomentum()
```

## Get Returns

```{r}
CSIQ_90d_return <- dailyReturn(CSIQ_90d)
dygraph(CSIQ_90d_return)
```

### Stats on returns

```{r}

summary(CSIQ_90d_return)
sum(CSIQ_90d_return)

ggplot(CSIQ_90d_return, aes(x=CSIQ_90d_return))+geom_histogram()+
  geom_vline(xintercept = median(CSIQ_90d_return))+
  geom_vline(xintercept = quantile(CSIQ_90d_return, 0.2))+
  geom_vline(xintercept = quantile(CSIQ_90d_return, 0.8))+
  scale_y_continuous(expand=c(0,0))+
  scale_x_continuous(labels=percent)+
  labs(title="Daily returns over the period", 
       subtitle='Lines: 20th percentile, Median, 80th percentile',
       x='Daily % chg')

```

## Time Series Manipulation

Leverage xts to work with time series objects (probably warrants its own file)

### End points by period

* day, week, month, qtr, yr.
* returns the locations of each last item in the specified type of period, aligned on calendar.
* will include the last item available in the last period (last data point).
* returns a vector starting with 0 and ending with the extent (last row) of your data.
* can also use k to skip: on='years', k=4 returns location of every 4th yr in data (plus last)

```{r}
## get index number of last day of each year
index_yr <- endpoints(CSIQ, on="years")
## apply to time series to extract selected end points
CSIQ_yr_end <- CSIQ[index_yr]
head(CSIQ_yr_end[,c(2,3,5)]) ## selected cols for space

```

```{r}
index_week_end <- endpoints(CSIQ_Aug, on="weeks")
CSIQ_Aug[index_week_end, 1:4] ## open, high, low, close selected
```

### Summarization

```{r}
##
period.apply(CSIQ_Aug, INDEX=index_week_end, FUN=mean)

```

### Shortcut summarization

```{r}
## apply.weekly, apply.monthly, apply.quarterly, apply.yearly to summarize
## FUN=<whatever summarization function for the period)
apply.weekly(CSIQ_Aug, FUN=mean)
```


### Split into periods

```{r}
## split data into weeks - chunk the data into 'week' components, without summarizing
CSIQ_Aug_weeks <- split(CSIQ_Aug, f='weeks')
## second week: high, low, close selected
CSIQ_Aug_weeks[[2]][,2:4]
```

## Aggregate time periods

* Aggregate data by period
* Can take a single variable and calculate OHLC statistics by period

```{r}
CSIQ_yearly <- to.period(CSIQ[,6], period='years', name='CSIQ_adj', OHLC=TRUE)
CSIQ_yearly
```

### Shortcut aggregation

```{r}
CSIQ_2020_mth_agg <- to.monthly(CSIQ_2020[,6], name='CSIQ_adj')
CSIQ_2020_mth_agg
```

