---
title: "Stock-performance-basic"
author: "John Yuill"
date: "August 29, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(lubridate)
library(scales)

library(xts)
library(zoo)
library(quantmod)
library(dygraphs)

## theme set for ggplot
theme_set(theme_light())
```

## Get price data

Getting daily price data for any stock is as easy as using quantmod package and getSymbols function.

```{r}
getSymbols("CSIQ")

```

Entire history of the daily stock price (at least going back to 2007) is now available in an xts object named after the stock symbol. 

Provides daily Open, High, Low, Close, Volume, Adjusted (aka OHLCVA):

```{r}
## show first rows 
head(CSIQ)
## show last rows
tail(CSIQ)
```

"Adjusted" = prices adjusted for stock splits and dividends (possibly other developments? major new issues? buy-backs?)

### Filter for time period

* When importing: specific period (start date and end date)

```{r}
getSymbols("CSIQ", from="2018-01-01", to=Sys.Date())
CSIQ_2018 <- CSIQ

head(CSIQ_2018)

## reset full CSIQ for later
getSymbols("CSIQ")
```

* filter after import

```{r}
CSIQ_2020 <- CSIQ['2020-01-01/'] ## from specified date fwd
CSIQ_90d <- tail(CSIQ, n=90) ## last n 

## selected period
CSIQ_Aug <- CSIQ['2020-08/2020-08-31'] ## Aug 1-31
```



## Visualize 

Dygraphs pkg enables interactive charts designed to work with quantmod right out of the boox.

### Basic line chart

* chart_Series 
* dygraph

```{r}
## basic chart_Series from quantmod
chart_Series(OHLC(CSIQ))
## use dygraphs pkg
dygraph(OHLC(CSIQ))

```

* with volume

```{r}
barChart(CSIQ)
```

* closing price only

```{r}
chart_Series(Cl(CSIQ))
```

```{r}

```


### Get multiple stocks 

Each imported as separate object

```{r}
getSymbols(c("NEP","SPWR"))
## OR
stocklist <- c("NEP","SPWR","FSLR")
getSymbols(stocklist)
```


Easily add annotations

```{r}
graph <- dygraph(OHLC(CSIQ), main="Cdn Solar")
graph <- dyEvent(graph, "2020-02-21","Start of covid crash",
                 labelLoc="top")
graph

```

Show only adjusted close, filtered for 2020

```{r}
CSIQ_2020 <- CSIQ['2020-01-01/']
graph <- dygraph(Ad(CSIQ_2020), main="Cdn Solar")
graph <- dyEvent(graph, "2020-02-21","Start of covid crash",
                 labelLoc="top")
graph

```

## Adjusted price

Adjusted price is supposed to adjust for stock splits, while other prices are actual on the trading day.

However...Tesla (TSLA) had a 5-for-1 stock split on Aug 3 dropping price from ~$2,000 to ~$400. Prices collected as of Sep 5, 2020 show post-split throughout.


```{r}
getSymbols("TSLA")

## filter for Aug, pre-split
TSLA_presplit <- TSLA['2020-08/2020-08-30']
tail(TSLA_presplit)

```

```{r}
#chart_Series(OHLC(TSLA))
dygraph(OHLC(TSLA))
```


## Candlestick

With selected period - in this case, based on recent (suitable up to ~30 days at a time).

```{r}
CSIQ_recent <- tail(CSIQ, n=30)
graph <- dygraph(OHLC(CSIQ_recent))
dyCandlestick(graph)

```

* Open is either bottom of green bar OR top of red bar
* High is highest point (either line or top of bar)
* Low is lowest point (either line or bottom of bar)
* Close is either top of green bar OR bottom of red bar

## Add Technical features

* MACD: moving average convergence divergence signals
* BBands: Bollinger bands
* other features are available in quantmod (not shown)

```{r}
chart_Series(OHLC(CSIQ_2020))
add_MACD() 
add_BBands()
```

## Get Returns

```{r}
CSIQ_90d_return <- dailyReturn(CSIQ_90d)
dygraph(CSIQ_90d_return)
```

### Stats on returns

```{r}

summary(CSIQ_90d_return)
sum(CSIQ_90d_return)

ggplot(CSIQ_90d_return, aes(x=CSIQ_90d_return))+geom_histogram()+
  geom_vline(xintercept = median(CSIQ_90d_return))+
  geom_vline(xintercept = quantile(CSIQ_90d_return, 0.2))+
  geom_vline(xintercept = quantile(CSIQ_90d_return, 0.8))+
  scale_y_continuous(expand=c(0,0))+
  scale_x_continuous(labels=percent)+
  labs(title="Daily returns over the period", 
       subtitle='Lines: 20th percentile, Median, 80th percentile',
       x='Daily % chg')

```

## aggregate time periods