---
title: "Crytpo Currency Exploration"
author: "JY"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(quantmod) ## includes zoo (which includes xts)
library(PerformanceAnalytics)
library(tidyverse)
library(lubridate)
library(scales)
library(dygraphs)

theme_set(theme_light())
```

## Get Bitcoin data

* Cdn dollar value

```{r}
btc <- getSymbols(Symbols="BTC-CAD", auto.assign=FALSE)
## remove today
btc <- btc[1:nrow(btc)-1,]
summary(btc)

```

### Take a look

```{r}
## basic chart_Series from quantmod
chart_Series(Cl(btc))
```

```{r}
btc45d <- tail(btc, n=45)
summary(btc45d[,1:4])
chart_Series(btc45d)
```


## Get Ethereum data

```{r}
eth <- getSymbols(Symbols="ETH-CAD", auto.assign=FALSE)
## remove today - empty
eth <- eth[1:nrow(eth)-1,]
summary(eth)
```

### Take a look

```{r}
## basic chart_Series from quantmod
chart_Series(Cl(eth))
```

### Zoom in recent periods

* 90d

```{r}
eth90d <- tail(eth, n=90)
summary(eth90d[,1:4])
```

```{r}
chart_Series(eth90d)
```

* 45d

```{r}
eth45d <- tail(eth, n=45)
summary(eth45d[,1:4])
chart_Series(eth45d)
```


## Correlation: BTC, ETH

* since beginning 2017

```{r}
kryptos <- merge(btc, eth, join='outer', fill=NA)
kryptos2017 <- kryptos['2017-01-01/'] 
```

```{r}
chart_Series(kryptos[,c(4,8)])
```

```{r}
kryptos_all <- kryptos
kryptos <- kryptos["2017-01-01/"]
chart.Correlation(kryptos[,c(4,10)], histogram=TRUE)
```

### Explore Correlation

```{r}
df_kryptos <- as.data.frame(kryptos)
df_kryptos$date <- row.names(df_kryptos)
df_kryptos %>% ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close))+geom_point()
```

```{r}
df_kryptos <- df_kryptos %>% mutate(
  yr=year(date),
  mth=month(date)
)
```

```{r}
df_kryptos %>% ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close, color=as.factor(yr)))+geom_point()
```

Correlations vary by year. Of most relevance in most recent yr correlation.

###  2020

```{r}
df_kryptos_2020 <- df_kryptos %>% filter(date>='2020-01-01' & date<='2020-12-31')
```


```{r}
df_kryptos_2020 %>%
  ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close))+geom_point()+
  geom_smooth(method='lm')
```

#### By Month

```{r}
df_kryptos_2020 %>%
  ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close, color=as.factor(mth)))+geom_point()+
  geom_smooth(method='lm')
```

```{r}
df_kryptos_2020 %>% 
  ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close))+geom_point(aes(color=as.factor(mth)))+
  geom_smooth(method='lm')
```

```{r}

chart.Correlation(df_kryptos_2020[,c(2,3,4,5,8,9,10,11)], histogram=TRUE)

```

### 2021

```{r}
df_kryptos_2021 <- df_kryptos %>% filter(date>='2021-01-01' & date<='2021-12-31')
```

```{r}
df_kryptos_2021 %>%
  ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close))+geom_point()+
  geom_smooth(method='lm')
```

#### By Month

```{r}
df_kryptos_2021 %>%
  ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close, color=as.factor(mth)))+geom_point()+
  geom_smooth(method='lm')
```

```{r}
df_kryptos_2021 %>% 
  ggplot(aes(x=BTC.CAD.Close, y=ETH.CAD.Close))+geom_point(aes(color=as.factor(mth)))+
  geom_smooth(method='lm')
```

## Distribution

### 2020

#### Prices 

##### BTC

```{r}
df_kryptos_2020 %>% ggplot(aes(x=as.factor(mth), y=BTC.CAD.Close))+geom_boxplot()
```

##### ETH

```{r}
df_kryptos_2020 %>% ggplot(aes(x=as.factor(mth), y=ETH.CAD.Close))+geom_boxplot()
```

#### Returns

##### BTC

```{r}
kryptos2020 <- kryptos_all["2020-01-01/2020-12-31"]
kryptos2020_btc_ret <- dailyReturn(kryptos2020$BTC.CAD.Close)
dygraph(kryptos2020_btc_ret)
```

```{r}
kryptos2020_btc_ret %>% ggplot(aes(x=.))+geom_histogram()+
  geom_vline(xintercept = mean(kryptos2020_btc_ret))

```

##### ETH

```{r}
kryptos2020 <- kryptos_all["2020-01-01/2020-12-31"]
kryptos2020_eth_ret <- dailyReturn(kryptos2020$ETH.CAD.Close)
dygraph(kryptos2020_eth_ret)
```
```{r}
kryptos2020_eth_ret %>% ggplot(aes(x=.))+geom_histogram()+
  geom_vline(xintercept = mean(kryptos2020_eth_ret))

```

##### Combined

```{r}
kryptos2020_ret <- merge(kryptos2020_btc_ret, kryptos2020_eth_ret, join='outer', fill=NA)

dygraph(kryptos2020_ret)
```

```{r}
## convert to data frame for scatterplots etc
df_kryptos_2020_ret <- as.data.frame(kryptos2020_ret)
colnames(df_kryptos_2020_ret) <- c("BTC_ret","ETH_ret")

```

```{r}
ggplot(df_kryptos_2020_ret, aes(x=BTC_ret, y=ETH_ret))+geom_point()+
  geom_smooth(method='lm')

df_kryptos_2020_ret %>% ggplot(aes(x=BTC_ret))+geom_histogram()+
  geom_vline(xintercept=sd(df_kryptos_2020_ret$BTC_ret), color='blue')+
  geom_vline(xintercept=mean(df_kryptos_2020_ret$BTC_ret), color='green')+
  geom_vline(xintercept=0-sd(df_kryptos_2020_ret$BTC_ret), color='blue')+
  geom_text(aes(label=round(sd(BTC_ret),3)), color='blue', 
                x=sd(df_kryptos_2020_ret$BTC_ret), y=70, hjust= -0.1)+
  geom_text(aes(label=round(mean(BTC_ret),3)), color='blue',
            x=mean(df_kryptos_2020_ret$BTC_ret), y=100, hjust= -0.2)

df_kryptos_2020_ret %>% ggplot(aes(x=ETH_ret))+geom_histogram()+
  geom_vline(xintercept=sd(df_kryptos_2020_ret$ETH_ret), color='blue')+
  geom_vline(xintercept=0-sd(df_kryptos_2020_ret$ETH_ret), color='blue')+
  geom_vline(xintercept=mean(df_kryptos_2020_ret$ETH_ret), color='green')+
  geom_text(aes(label=round(sd(ETH_ret),3)), color='blue', 
                x=sd(df_kryptos_2020_ret$ETH_ret), y=70, hjust= -0.1)+
  geom_text(aes(label=round(mean(ETH_ret),3)), color='green',
            x=mean(df_kryptos_2020_ret$ETH_ret), y=100, hjust= -0.2)

```

### 2021

#### Prices 

##### BTC

```{r}
df_kryptos_2021 %>% ggplot(aes(x=as.factor(mth), y=BTC.CAD.Close))+geom_boxplot()
```

##### ETH

```{r}
df_kryptos_2021 %>% ggplot(aes(x=as.factor(mth), y=ETH.CAD.Close))+geom_boxplot()
```

#### Returns

##### BTC

```{r}
kryptos2021 <- kryptos_all["2021-01-01/2021-12-31"]
kryptos2021_btc_ret <- dailyReturn(kryptos2021$BTC.CAD.Close)
dygraph(kryptos2021_btc_ret)
```
```{r}
kryptos2021_btc_ret %>% ggplot(aes(x=.))+geom_histogram()+
  geom_vline(xintercept = mean(kryptos2021_btc_ret))

```

##### ETH

```{r}
kryptos2021 <- kryptos_all["2021-01-01/2021-12-31"]
kryptos2021_eth_ret <- dailyReturn(kryptos2021$ETH.CAD.Close)
dygraph(kryptos2021_eth_ret)
```

```{r}
kryptos2021_eth_ret %>% ggplot(aes(x=.))+geom_histogram()+
  geom_vline(xintercept = mean(kryptos2021_eth_ret))

```


##### Combined

```{r}
kryptos2021_ret <- merge(kryptos2021_btc_ret, kryptos2021_eth_ret, join='outer', fill=NA)

dygraph(kryptos2021_ret)
```

```{r}
## convert to data frame for scatterplots etc
df_kryptos_2021_ret <- as.data.frame(kryptos2021_ret)
colnames(df_kryptos_2021_ret) <- c("BTC_ret","ETH_ret")

```

```{r}
ggplot(df_kryptos_2021_ret, aes(x=BTC_ret, y=ETH_ret))+geom_point()+
  geom_smooth(method='lm')

df_kryptos_2021_ret %>% ggplot(aes(x=BTC_ret))+geom_histogram()+
  geom_vline(xintercept=sd(df_kryptos_2021_ret$BTC_ret), color='blue')+
  geom_vline(xintercept=mean(df_kryptos_2021_ret$BTC_ret), color='green')+
  geom_vline(xintercept=0-sd(df_kryptos_2021_ret$BTC_ret), color='blue')+
  geom_text(aes(label=round(sd(BTC_ret),3)), color='blue', 
                x=sd(df_kryptos_2021_ret$BTC_ret), y=70, hjust= -0.1)+
  geom_text(aes(label=round(mean(BTC_ret),3)), color='blue',
            x=mean(df_kryptos_2021_ret$BTC_ret), y=100, hjust= -0.2)

df_kryptos_2021_ret %>% ggplot(aes(x=ETH_ret))+geom_histogram()+
  geom_vline(xintercept=sd(df_kryptos_2021_ret$ETH_ret), color='blue')+
  geom_vline(xintercept=0-sd(df_kryptos_2021_ret$ETH_ret), color='blue')+
  geom_vline(xintercept=mean(df_kryptos_2021_ret$ETH_ret), color='green')+
  geom_text(aes(label=round(sd(ETH_ret),3)), color='blue', 
                x=sd(df_kryptos_2021_ret$ETH_ret), y=70, hjust= -0.1)+
  geom_text(aes(label=round(mean(ETH_ret),3)), color='green',
            x=mean(df_kryptos_2021_ret$ETH_ret), y=100, hjust= -0.2)

```

#### PERIOD RETURNS

At the end of the day...what matters is what your return is at the end of the day. :)

```{r}
df_kryptos_2021ytd <- df_kryptos_2021 %>% filter(date==min(date) | date==max(date)) %>% select(date, yr, mth, BTC.CAD.Close, ETH.CAD.Close) %>%
  mutate(BTC.CAD.ret=BTC.CAD.Close/lag(BTC.CAD.Close)-1,
         ETH.CAD.ret=ETH.CAD.Close/lag(ETH.CAD.Close)-1)

df_kryptos_2021ytd_lg <- df_kryptos_2021ytd %>% 
                          filter(date==max(date)) %>%
                          select(date, yr, mth, BTC.CAD.ret, ETH.CAD.ret) %>%
                          pivot_longer(cols=c(4:5),
                                      names_to='coin',
                                      values_to='ytd_return')

df_kryptos_2021ytd_lg %>% ggplot(aes(x=coin, y=ytd_return))+
                               geom_col()+
                              scale_y_continuous(labels=percent)

df_kryptos_2021ytd[,c(4:7)]

```

## MULTIPLE Currencies

### Get data

```{r}
ccurrency <- c("BTC-CAD", "ETH-CAD","LTC-CAD","BCH-CAD")

kryptos_multi <- NULL
for(c in ccurrency){
  kryptos_multi <- cbind(kryptos_multi,
                         getSymbols(Symbols=c, 
                                 from='2018-01-01', periodicity='daily', 
                                 auto.assign = FALSE)[,4])
}

```

### Check correlations

```{r}
chart.Correlation(kryptos_multi, histogram=TRUE)
```

