---
title: "Crypto Transactions & Returns"
author: "John Yuill"
date: "`r Sys.Date()`"
output: html_document
---

## Intro

This report compiles info on cryptocurrency transactions - buying and selling - and calculates returns on those transactions.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE)

library(quantmod) ## includes zoo (which includes xts)
library(PerformanceAnalytics)
library(tidyverse)
library(lubridate)
library(scales)
library(dygraphs)
library(here)
library(viridis)
library(gridExtra)

theme_set(theme_light())
```

```{r, IMPORT}
## IMPORT trade and cost data
## - processed in google sheets from CB export
## - imported, cleaned, saved via trades_data_import.R
trade_data <- read_csv(here('data/trades_data_exp.csv'))

## filter for crypto
#crypto_trades <- trade_data %>% filter(str_detect(Symbol, 'BTC|ETH|LTC'))

```

```{r, PROCESS}
## filter for SELL transactions
trade_sell <- trade_data %>% filter(Activity=='trade-sell') %>% select(-`Receipt No`,-`Date_time`
) %>% mutate(
  gain=ifelse(Currency=='BTC', `CAD Amount`+`BTC-cost`,
              ifelse(Currency=='ETH', `CAD Amount`+`ETH-cost`, `CAD Amount`+`LTC-cost`)),
  cost=`CAD Amount`-gain,
  avg_cost=ifelse(Currency=='BTC', `BTC-cost`/BTC,
              ifelse(Currency=='ETH', `ETH-cost`/ETH, `LTC-cost`/LTC)),
  roi=gain/cost
)

## summarize by currency
trade_sell_curr <- trade_sell %>% group_by(Currency) %>% summarize(
  units=sum(Amount),
  ttl_revenue=sum(`CAD Amount`),
  ttl_gain=sum(gain),
  ttl_cost=sum(cost),
  avg_price=ttl_revenue/units,
  avg_cost=ttl_cost/units
) %>% mutate(
  roi=ttl_gain/ttl_cost
)

## summarize all
trade_sell_all <- trade_sell %>% summarize(
  ttl_revenue=sum(`CAD Amount`),
  ttl_gain=sum(gain),
  ttl_cost=sum(cost)
) %>% mutate(
  roi=ttl_gain/ttl_cost
)

```

## Sell Transaction Returns

Performance info based on actual sales made.

### Overall Performance 

```{r, fig.height=3, fig.width=6}
psa <- trade_sell_all %>% ggplot(aes(x=1, y=ttl_gain, label=prettyNum(round(ttl_gain,0), big.mark = ",")))+geom_col()+
  geom_text(size=8, nudge_y = 1200)+
  geom_hline(yintercept=20000)+
  scale_y_continuous(labels=dollar, expand = expansion(mult=c(0,0.05)))+
  labs(x='ttl gains v target')+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

ptr <- ggplot(trade_sell_curr, aes(x=1, y=ttl_gain, fill=Currency))+geom_col()+
  scale_y_continuous(labels=dollar, expand = expansion(mult=c(0,0.05)))+
  labs(x='ttl gains by currency')+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

proi <- ggplot(trade_sell_curr, aes(x=Currency, y=roi))+geom_col()+
  scale_y_continuous(labels=percent, expand = expansion(mult=c(0,0.05)))+
  labs(x="")

grid.arrange(psa, ptr, proi, nrow=1)
```


```{r}
total_invest <- trade_data %>% filter(Activity=='fund-e-transfer') %>% summarize(
  ttl_invest=sum(Amount)
)

#total_invest
```
Total invested to date: `r total_invest$ttl_invest`

### Individual Sell Transactions

```{r, fig.height=3, fig.width=8}

p_amt <- trade_sell %>% ggplot(aes(x=Date, y=gain, fill=Currency))+geom_col(position=position_dodge(preserve='single'))+
  scale_y_continuous(labels=dollar, expand=expansion(add=c(0,100)))+
  labs(x="")+theme(legend.position='none')

p_pc <- trade_sell %>% ggplot(aes(x=Date, y=roi, fill=Currency))+geom_col(position=position_dodge(preserve="single"))+
  scale_y_continuous(labels=percent, expand=expansion(add=c(0,0.1)))+
  labs(x="")

# ## VIZ
# p_amt <- ggplot(trade_dt_summary, aes(x=Date, y=total_gain, fill=Symbol))+geom_col()+
#   #coord_flip()+
#   scale_x_date()+
#   scale_y_continuous(labels=dollar, expand=c(0,0))+
#   scale_fill_viridis_d()+
#   labs(x='')+theme(legend.position='none')
#   
# 
# p_pc <- ggplot(trade_dt_summary, aes(x=Date, y=roi, fill=Symbol))+geom_col()+
#   #coord_flip()+
#   scale_x_date()+
#   scale_y_continuous(labels=percent, expand=c(0,0))+
#   scale_fill_viridis_d()+
#   labs(x='')
# 
grid.arrange(p_amt, p_pc, nrow=1)
```

## Potential Returns

Potential returns based on:

* Current prices for various currencies
* Latest average cost for various currencies

```{r}
## get current prices

```

## Value at Risk

Value at Risk for each currency calculated based on:

* 5% likelihood drop in price over 24hrs
* amount invested

```{r}

```

