---
title: "Krytpo Porto Status"
author: "JY"
date: "`r Sys.Date()`"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE)

library(quantmod) ## includes zoo (which includes xts)
library(PerformanceAnalytics)
library(tidyverse)
library(lubridate)
library(scales)
library(dygraphs)
library(googlesheets4)
library(here)
library(viridis)
library(RColorBrewer)
library(kableExtra)
library(formattable)
library(gridExtra)
library(plotly)

theme_set(theme_light())
options(scipen = 20)

## colors
price_line <- 'blue4'
profit <- 'darkgreen'
loss <- 'darkred'

```

## Purpose 

Keep track of status of krypto portfolio over time, based on transactions combined with current price data.

* Import krypto transaction data from spreadsheet.
* Download latest price data.
* Combine to calculate theoretical profit/loss trends.
    + by currency
    + overall portfolio
    
## Method {.tabset}

1. Log into crypto account.
2. Download activity report.
3. Go to spreadsheet in Marengo - check most recent transaction date there.
4. Copy/paste new transactions from activity report.
5. Copy down formulas in spreadsheet.

```{r}
gcred <- 'original-return-107905-8542c2ab0ad1.json'
if(file.exists(here(gcred))){
  gs4_auth(path=here(gcred))
}
gsheet <- 'https://docs.google.com/spreadsheets/d/1DRW5n6s6UtDAq3-kg-JTD4Yd2_24QfjAAxSU7-itkEE/edit?usp=sharing'
sheet_item <- 'krypton-exp'

ktrans <- read_sheet(ss=gsheet, sheet=sheet_item, skip=1)
ktrans <- ktrans[,1:7]

## set up table to hold latest profit data
curr_profits <- data.frame()
```

```{r FUNCTIONS}

## calculate average cost base
acb <- function(currency){
  curr_trans <- ktrans %>% filter(Currency==currency)
  curr_trans$Date <- ymd_hms(curr_trans$Date)
  curr_trans$date_trans <- date(curr_trans$Date)
  
  ## calc average cost (adjusted cost base)
## 2 aspects:
## - 1. basic: cumulative amount spent / cumulative units acquired (weighted average)
## - 2. account for sales: need to remove the units sold from cumulative total and 
##   reduce amount spent by number of units sold at a value of previous acb
## RETURNS curr_trans 
curr_trans <- curr_trans %>% 
  mutate(
    units_calc=ifelse(Activity=='trade-buy',Amount,Amount* -1),
    cumunits_calc=cumsum(units_calc),
    cost_calc=ifelse(Activity=='trade-buy',`CAD Amount`,0),
    cost_sell_calc=ifelse(Activity=='trade-sell',units_calc*(cumsum(cost_calc)/lag(cumunits_calc)),0),
    cumcost_calc=ifelse(Activity=='trade-buy',cumsum(cost_calc)+cumsum(cost_sell_calc),cumsum(cost_calc)-abs(units_calc)*(cumsum(cost_calc)/lag(cumunits_calc))),
    acb=cumcost_calc/cumunits_calc
  )
  return(curr_trans)
}

## generate transaction charts
## RETURNS 3 charts: individual transactions and cumulative
curr_trans_chart <- function(){
  chart_title <- "Transactions - Date, Price, Rel. Size"
  ct0 <- curr_trans %>% ggplot(aes(x=date_trans, y=`CAD Rate`, size=Amount, color=Activity))+geom_point()+
    scale_y_continuous(labels=dollar)+
    labs(title=chart_title, x="", y="Price")
  chart_title_01 <- "Ave Cost at Each Transaction"
  ct1 <- curr_trans %>% ggplot(aes(x=date_trans, y=acb))+geom_col(position=position_dodge2())+
    scale_y_continuous(labels=dollar)+
    labs(x="")
  chart_title_02 <- "Cumulative Units at Each Transaction"
  ct2 <- curr_trans %>% ggplot(aes(x=date_trans, y=cumunits_calc))+geom_col(position=position_dodge2())+
    labs(title=paste(chart_title_01, " - ",chart_title_02), x="")
  return(list(ct0, ct1, ct2))
}

### PRICE DATA
## RETURNS 3 data frames: curr price cost for currency, curr profits for currency, combined profits for currencies
price <- function(currency, symbol){
  symbol_price <- getSymbols(Symbols=symbol, auto.assign=FALSE)
  ## match dates with cost data
  dt_start <- as.character(min(curr_trans$date_trans))
  ## GENERIC object name for easy re-use
  symbol_close <- paste0(symbol,".Close")
  curr_price <- symbol_price[paste0(dt_start,"/"),symbol_close]
  ## convert ta data frame and combine with cost data
  curr_price_df <- as.data.frame(curr_price)
  curr_price_df$date <- date(row.names(curr_price_df))
  row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
  curr_price_df <- curr_price_df %>% select(date, symbol_close)
  
  ## COMBINE with cost data
  ## get essential cols from trans 
  curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
  ## need to have only one entry per date -> if there are multiple, should be last item for any date
  curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>% 
    summarize(cumunits_calc=last(cumunits_calc),
              cumcost_calc=last(cumcost_calc), 
              acb=last(acb)
            ) %>%
    rename(date=date_trans)
  
  ## join price and cost data
  curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
  ## fill in NA values with previous value
  curr_price_cost <- na.locf(curr_price_cost) 
  
  ## CALC PROFIT on any given day
  curr_price_cost <- curr_price_cost %>% mutate(
    ## use col reference so works regardless of col name (which will be based on currency)
    market_val=curr_price_cost[[2]]*cumunits_calc,
    book_val=cumunits_calc*acb,
    profit=market_val-book_val
  )
  
  ## CALC PROFIT AND ROI to date - compile for currencies
  curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
    mutate(currency=currency,
           roi=market_val/book_val-1)
  curr_profits <- bind_rows(curr_profits, curr_profit)
  
  return(list(curr_price_cost, curr_profit, curr_profits))
}

## Creates price-cost timeline charts
## RETURNS: two charts - timeline price vs cost, timeline net return
price_cost_charts <- function(curr_price_cost, symbol){
  ## Price v Cost charts
  chart_title_01 <- paste0(symbol,": Price vs Cost")
  symbol_close <- paste0(symbol,".Close") ## assemble variable name then use .data[[variable name]]
  pcp1 <- curr_price_cost %>% ggplot(aes(x=date, y=.data[[symbol_close]]))+geom_line(color=price_line)+
    geom_line(aes(x=date, y=acb), color='black')+
    scale_y_continuous(labels=dollar)+
    labs(title=chart_title_01, x="")
  
  ## Profit
  chart_title_02 <- paste0(symbol,": Net Profit")
  pcp2 <- curr_price_cost %>% ggplot(aes(x=date, y=profit))+geom_col(aes(fill=profit>0))+ ## color condition
    scale_y_continuous(labels=dollar)+
    scale_fill_manual(values = setNames(c(profit,loss),c(T, F)))+ ## manual color based on condition
    theme(legend.position = "none")+ ## hide legend - not needed
    labs(title=paste(chart_title_01, " - ",chart_title_02), x="")
  
  #grid.arrange(pcp1, pcp2, nrow=1)
  ## plotly version - using subplot
  pcp1_p <- ggplotly(pcp1)
  pcp2_p <- ggplotly(pcp2)
  return(list(pcp1_p, pcp2_p))
}

```


### Bitcoin

#### Transactions & Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
currency <- 'BTC'
ct <- acb(currency) ## function for ave cost data
curr_trans <- ct ## rename to match existing code
ct_chart <- curr_trans_chart() ## get charts
ggplotly(ct_chart[[1]])
subplot(ggplotly(ct_chart[[2]]), ggplotly(ct_chart[[3]]), nrows=1)

## SAVE (session) for currency
btc_trans_save <- curr_trans

```

#### Prices, Costs, Profits

* Cdn dollar value

1. Get latest price data.
2. Combine with cost data.
3. Calculate profit and ROI.

```{r}
### PRICE DATA
# btc <- getSymbols(Symbols="BTC-CAD", auto.assign=FALSE)
# ## remove today - or not, in order to get latest price
# #btc <- btc[1:nrow(btc)-1,]
# 
# ## match dates with cost data
# dt_start <- as.character(min(curr_trans$date_trans))
# 
# ## GENERIC object name for easy re-use
# curr_price <- btc[paste0(dt_start,"/"),"BTC-CAD.Close"]
# 
# ## convert ta data frame and combine with cost data
# curr_price_df <- as.data.frame(curr_price)
# curr_price_df$date <- date(row.names(curr_price_df))
# row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
# curr_price_df <- curr_price_df %>% select(date, `BTC-CAD.Close`)
# 
# ## COMBINE with cost data
# ## get essential cols from trans 
# curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
# ## need to have only one entry per date -> if there are multiple, should be last item for any date
# curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>% summarize(cumunits_calc=last(cumunits_calc),
#                                                                         cumcost_calc=last(cumcost_calc),
#                                                                         acb=last(acb)
#                                                                         ) %>%
#   rename(date=date_trans)
# 
# ## join price and cost data
# curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
# ## fill in NA values with previous value
# curr_price_cost <- na.locf(curr_price_cost)
# 
# ## CALC PROFIT on any given day
# curr_price_cost <- curr_price_cost %>% mutate(
#   market_val=`BTC-CAD.Close`*cumunits_calc,
#   book_val=cumunits_calc*acb,
#   profit=market_val-book_val
# )
# 
# ## CALC PROFIT AND ROI to date - compile for currencies
# curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
#   mutate(currency=currency,
#          roi=market_val/book_val-1)
# curr_profits <- bind_rows(curr_profits, curr_profit)

## ABOVE REPLACED BY FUNCTION ====
symbol <- 'BTC-CAD'
price_cost_data <- price(currency, symbol) 
curr_price_cost <- price_cost_data[[1]]
curr_profits <- price_cost_data[[3]]
```

```{r, fig.height=4, fig.width=9}
## Price v Cost charts
# chart_title <- paste0(symbol,": Price vs Cost")
# pcp1 <- curr_price_cost %>% ggplot(aes(x=date, y=`BTC-CAD.Close`))+geom_line(color=price_line)+
#   geom_line(aes(x=date, y=acb), color='black')+
#   scale_y_continuous(labels=dollar)+
#   labs(title=chart_title, x="")
# 
# ## Profit
# chart_title <- paste0(symbol,": Net Profit")
# pcp2 <- curr_price_cost %>% ggplot(aes(x=date, y=profit))+geom_col(aes(fill=profit>0))+ ## color condition
#   scale_y_continuous(labels=dollar)+
#   scale_fill_manual(values = setNames(c(profit,loss),c(T, F)))+ ## manual color based on condition
#   theme(legend.position = "none")+ ## hide legend - not needed
#   labs(title=chart_title, x="")
# 
# #grid.arrange(pcp1, pcp2, nrow=1)
# ## plotly version - using subplot
# pcp1_p <- ggplotly(pcp1)
# pcp2_p <- ggplotly(pcp2)
# subplot(pcp1_p, pcp2_p, nrows=1)

## ALL ABOVE REPLACED BY FUNCTION ====
pc_charts <- price_cost_charts(curr_price_cost, symbol)
subplot(ggplotly(pc_charts[[1]]), ggplotly(pc_charts[[2]]), nrows=1)
```

### Ethereum

#### Transactions & Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## ethereum
currency <- 'ETH'
ct <- acb(currency) ## function for ave cost data
curr_trans <- ct ## rename to match existing code
ct_chart <- curr_trans_chart() ## get charts
ggplotly(ct_chart[[1]])
subplot(ggplotly(ct_chart[[2]]), ggplotly(ct_chart[[3]]), nrows=1)

## SAVE (session) for currency
eth_trans_save <- curr_trans
```

#### Price, Costs, Profits

```{r}
### Price Data
symbol <- "ETH-CAD" ## used in getSymbols call and chart titles
price_cost_data <- price(currency, symbol) 
curr_price_cost <- price_cost_data[[1]]
curr_profits <- price_cost_data[[3]]
```

```{r, fig.height=4, fig.width=9}
### PRICE V COST charts
pc_charts <- price_cost_charts(curr_price_cost, symbol)
subplot(ggplotly(pc_charts[[1]]), ggplotly(pc_charts[[2]]), nrows=1)
```

### Cardano

#### Transactions & Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## cardano
currency <- 'ADA'
ct <- acb(currency) ## function for ave cost data
curr_trans <- ct ## rename to match existing code
ct_chart <- curr_trans_chart() ## get charts
ggplotly(ct_chart[[1]])
subplot(ggplotly(ct_chart[[2]]), ggplotly(ct_chart[[3]]), nrows=1)

## SAVE (session) for currency
ada_trans_save <- curr_trans
```

#### Prices, Costs, Profits

```{r}
### PRICE DATA
symbol <- "ADA-CAD" ## used in getSymbols call and chart titles
#price_col <- "ADA-CAD.Close" ## not used because col automatically identified in function - could be used later
price_cost_data <- price(currency, symbol) 
curr_price_cost <- price_cost_data[[1]]
curr_profits <- price_cost_data[[3]]
```

```{r, fig.height=4, fig.width=9}
### PRICE V COST charts
pc_charts <- price_cost_charts(curr_price_cost, symbol)
subplot(ggplotly(pc_charts[[1]]), ggplotly(pc_charts[[2]]), nrows=1)
```

### Polkadot

#### Transactions & Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## polkadot
currency <- 'DOT'
ct <- acb(currency) ## function for ave cost data
curr_trans <- ct ## rename to match existing code
ct_chart <- curr_trans_chart() ## get charts
ggplotly(ct_chart[[1]])
subplot(ggplotly(ct_chart[[2]]), ggplotly(ct_chart[[3]]), nrows=1)

## SAVE (session) for currency
dot_trans_save <- curr_trans

```

#### Prices, Costs, Profits

```{r}
### PRICE DATA
symbol <- "DOT1-CAD" ## used in getSymbols call and chart titles
price_cost_data <- price(currency, symbol) 
curr_price_cost <- price_cost_data[[1]]
curr_profits <- price_cost_data[[3]]
```

```{r, fig.height=4, fig.width=9}
### PRICE V COST charts
pc_charts <- price_cost_charts(curr_price_cost, symbol)
subplot(ggplotly(pc_charts[[1]]), ggplotly(pc_charts[[2]]), nrows=1)
```

### Litecoin

#### Transactions & Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## litecoin
currency <- 'LTC'
ct <- acb(currency) ## function for ave cost data
curr_trans <- ct ## rename to match existing code
ct_chart <- curr_trans_chart() ## get charts
ggplotly(ct_chart[[1]])
subplot(ggplotly(ct_chart[[2]]), ggplotly(ct_chart[[3]]), nrows=1)

## SAVE (session) for currency
ltc_trans_save <- curr_trans

```

#### Prices, Costs, Profits

```{r}
### PRICE DATA
symbol <- "LTC-CAD" ## used in getSymbols call and chart titles
price_cost_data <- price(currency, symbol) 
curr_price_cost <- price_cost_data[[1]]
curr_profits <- price_cost_data[[3]]
```


```{r, fig.height=4, fig.width=9}
### PRICE V COST charts
pc_charts <- price_cost_charts(curr_price_cost, symbol)
subplot(ggplotly(pc_charts[[1]]), ggplotly(pc_charts[[2]]), nrows=1)
```

### TOTALS

```{r, fig.height=3.5, fig.width=9}
#kable(curr_profits)
curr_profits <- curr_profits %>% select(date, currency, book_val, market_val, profit, roi) %>% mutate(
  book_val=round(book_val, 0),
  market_val=round(market_val, 0),
  profit=round(profit,0),
  roi=round(roi, 2)
)

chart_title <- "Current Profit Standing by Currency"
pcp <- curr_profits %>% ggplot(aes(x=currency, y=profit))+geom_col()+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title, x="")

chart_title <- "Current ROI by Currency"  
pcr <- curr_profits %>% ggplot(aes(x=currency, y=roi))+geom_col()+
  scale_y_continuous(labels=percent)+
  labs(title=chart_title, x="")

grid.arrange(pcp, pcr, nrow=1)
```

```{r}
## calc total performance
total_perform <- curr_profits %>% summarize(
  date=max(date),
  total_invested=sum(book_val),
  total_market_val=sum(market_val),
  total_profit=sum(profit),
  total_roi=sum(market_val)/sum(book_val)-1)

## functions for formattable
currency_format <- formatter("span", x ~ currency(x, digits=0))
acct_format <- formatter("span", x ~ accounting(x, digits=0, big.mark=","))
pc_format <- formatter("span", x ~ percent(x, digits=0))

formattable(curr_profits, list(
  book_val=currency_format,
  market_val=currency_format,
  profit=acct_format,
  roi=pc_format))

#curr_profits %>% kbl() %>% kable_styling()
formattable(total_perform, list(
  total_invested=currency_format,
  total_market_val=currency_format,
  total_profit=currency_format,
  total_roi=pc_format))

```

