---
title: "Krytpo Porto Status"
author: "JY"
date: "`r Sys.Date()`"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message=FALSE,
                      warning=FALSE)

library(quantmod) ## includes zoo (which includes xts)
library(PerformanceAnalytics)
library(tidyverse)
library(lubridate)
library(scales)
library(dygraphs)
library(googlesheets4)
library(here)
library(viridis)
library(RColorBrewer)
library(kableExtra)
library(formattable)
library(gridExtra)
library(plotly)

theme_set(theme_light())
options(scipen = 20)

## colors
price_line <- 'blue4'
profit <- 'darkgreen'
loss <- 'darkred'

```

## Purpose {.tabset}

Keep track of status of krypto portfolio over time, based on transactions combined with current price data.

* Import krypto transaction data from spreadsheet.
* Download latest price data.
* Combine to calculate theoretical profit/loss trends.
    + by currency
    + overall portfolio
    
## Method

1. Log into crypto account.
2. Download activity report.
3. Go to spreadsheet in Marengo - check most recent transaction date there.
4. Copy/paste new transactions from activity report.
5. Copy down formulas in spreadsheet.

```{r}
gcred <- 'original-return-107905-8542c2ab0ad1.json'
if(file.exists(here(gcred))){
  gs4_auth(path=here(gcred))
}
gsheet <- 'https://docs.google.com/spreadsheets/d/1DRW5n6s6UtDAq3-kg-JTD4Yd2_24QfjAAxSU7-itkEE/edit?usp=sharing'
sheet_item <- 'krypton-exp'

ktrans <- read_sheet(ss=gsheet, sheet=sheet_item, skip=1)
ktrans <- ktrans[,1:7]

## set up table to hold latest profit data
curr_profits <- data.frame()
```

```{r FUNCTIONS}

## calculate average cost base
acb <- function(currency){
  curr_trans <- ktrans %>% filter(Currency==currency)
  curr_trans$Date <- ymd_hms(curr_trans$Date)
  curr_trans$date_trans <- date(curr_trans$Date)
  
  ## calc average cost (adjusted cost base)
## 2 aspects:
## - 1. basic: cumulative amount spent / cumulative units acquired (weighted average)
## - 2. account for sales: need to remove the units sold from cumulative total and 
##   reduce amount spent by number of units sold at a value of previous acb
curr_trans <- curr_trans %>% 
  mutate(
    units_calc=ifelse(Activity=='trade-buy',Amount,Amount* -1),
    cumunits_calc=cumsum(units_calc),
    cost_calc=ifelse(Activity=='trade-buy',`CAD Amount`,0),
    cost_sell_calc=ifelse(Activity=='trade-sell',units_calc*(cumsum(cost_calc)/lag(cumunits_calc)),0),
    cumcost_calc=ifelse(Activity=='trade-buy',cumsum(cost_calc)+cumsum(cost_sell_calc),cumsum(cost_calc)-abs(units_calc)*(cumsum(cost_calc)/lag(cumunits_calc))),
    acb=cumcost_calc/cumunits_calc
  )
  return(curr_trans)
}

## generate transaction charts
curr_trans_chart <- function(){
  ct1 <- curr_trans %>% ggplot(aes(x=date_trans, y=acb))+geom_col()+
    scale_y_continuous(labels=dollar)+
    labs(x="")
  ct2 <- curr_trans %>% ggplot(aes(x=date_trans, y=cumunits_calc))+geom_col()+
    labs(x="")
  return(list(ct1, ct2))
}

### PRICE DATA
price <- function(currency, symbol){
  symbol_price <- getSymbols(Symbols=symbol, auto.assign=FALSE)
  ## match dates with cost data
  dt_start <- as.character(min(curr_trans$date_trans))
  ## GENERIC object name for easy re-use
  symbol_close <- paste0(symbol,".Close")
  curr_price <- symbol_price[paste0(dt_start,"/"),symbol_close]
  ## convert ta data frame and combine with cost data
  curr_price_df <- as.data.frame(curr_price)
  curr_price_df$date <- date(row.names(curr_price_df))
  row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
  curr_price_df <- curr_price_df %>% select(date, symbol_close)
 
  ## COMBINE with cost data
  ## get essential cols from trans 
  curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
  ## need to have only one entry per date -> if there are multiple, should be last item for any date
  curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>%   summarize(cumunits_calc=last(cumunits_calc),
                                                                        cumcost_calc=last(cumcost_calc),
                                                                        acb=last(acb)
                                                                        ) %>%
  rename(date=date_trans)

  ## join price and cost data
  curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
  ## fill in NA values with previous value
  curr_price_cost <- na.locf(curr_price_cost) 
  
  ## CALC PROFIT on any given day
curr_price_cost <- curr_price_cost %>% mutate(
  ## use col reference so works regardless of col name (which will be based on currency)
  market_val=curr_price_cost[[2]]*cumunits_calc,
  book_val=cumunits_calc*acb,
  profit=market_val-book_val
)

## CALC PROFIT AND ROI to date - compile for currencies
curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
  mutate(currency=currency,
         roi=market_val/book_val-1)
curr_profits <- bind_rows(curr_profits, curr_profit)

}




##======================
## TEST AREA
ct <- acb('BTC')
curr_trans <- ct
ct_chart <- curr_trans_chart()

#subplot(ggplotly(ct_chart[[1]]), ggplotly(ct_chart[[2]]), nrows=1)

price_data <- price('BTC','BTC-CAD') 
```


### Bitcoin

#### Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## bitcoin
currency <- 'BTC'
curr_trans <- ktrans %>% filter(Currency==currency)
curr_trans$Date <- ymd_hms(curr_trans$Date) ## format timestamp
curr_trans$date_trans <- date(curr_trans$Date) ## add date field for transaction

## calc average cost (adjusted cost base)
## 2 aspects:
## - 1. basic: cumulative amount spent / cumulative units acquired (weighted average)
## - 2. account for sales: need to remove the units sold from cumulative total and 
##   reduce amount spent by number of units sold at a value of previous acb
curr_trans <- curr_trans %>% 
  mutate(
    units_calc=ifelse(Activity=='trade-buy',Amount,Amount* -1),
    cumunits_calc=cumsum(units_calc),
    cost_calc=ifelse(Activity=='trade-buy',`CAD Amount`,0),
    cost_sell_calc=ifelse(Activity=='trade-sell',units_calc*(cumsum(cost_calc)/lag(cumunits_calc)),0),
    #cumcost=ifelse(Activity=='trade-buy',cumsum(cost),abs(units)*(cumsum(cost)/cumunits)),
    #cumcostunit=ifelse(Activity=='trade-buy',cumsum(units),abs(units)),
    #cumcostunitlag=ifelse(Activity=='trade-buy',0,lag(cumunits)),
    #cumcostcost=ifelse(Activity=='trade-buy',0,cumsum(cost)),
    #cumcostcum=ifelse(Activity=='trade-buy',0,cumsum(cost)/lag(cumunits)),
    #cumcostsell=ifelse(Activity=='trade-buy',0,units*(cumsum(cost)/lag(cumunits))),
    cumcost_calc=ifelse(Activity=='trade-buy',cumsum(cost_calc)+cumsum(cost_sell_calc),cumsum(cost_calc)-abs(units_calc)*(cumsum(cost_calc)/lag(cumunits_calc))),
    acb=cumcost_calc/cumunits_calc
  )

ct1 <- curr_trans %>% ggplot(aes(x=date_trans, y=acb))+geom_col()+
  scale_y_continuous(labels=dollar)+
  labs(x="")
ct2 <- curr_trans %>% ggplot(aes(x=date_trans, y=cumunits_calc))+geom_col()+
  labs(x="")

#grid.arrange(ct1, ct2, nrow=1)
## subplot instead of grid.arrange for plotly charts
subplot(ggplotly(ct1), ggplotly(ct2), nrows=1)

## SAVE (session) for currency
btc_trans_save <- curr_trans

```

#### Prices, Costs, Profits

* Cdn dollar value

1. Get latest price data.
2. Combine with cost data.
3. Calculate profit and ROI.

```{r}
### PRICE DATA
btc <- getSymbols(Symbols="BTC-CAD", auto.assign=FALSE)
## remove today - or not, in order to get latest price
#btc <- btc[1:nrow(btc)-1,]

## match dates with cost data
dt_start <- as.character(min(curr_trans$date_trans))

## GENERIC object name for easy re-use
curr_price <- btc[paste0(dt_start,"/"),"BTC-CAD.Close"]

## convert ta data frame and combine with cost data
curr_price_df <- as.data.frame(curr_price)
curr_price_df$date <- date(row.names(curr_price_df))
row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
curr_price_df <- curr_price_df %>% select(date, `BTC-CAD.Close`)

## COMBINE with cost data
## get essential cols from trans 
curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
## need to have only one entry per date -> if there are multiple, should be last item for any date
curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>% summarize(cumunits_calc=last(cumunits_calc),
                                                                        cumcost_calc=last(cumcost_calc),
                                                                        acb=last(acb)
                                                                        ) %>%
  rename(date=date_trans)

## join price and cost data
curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
## fill in NA values with previous value
curr_price_cost <- na.locf(curr_price_cost)

## CALC PROFIT on any given day
curr_price_cost <- curr_price_cost %>% mutate(
  market_val=`BTC-CAD.Close`*cumunits_calc,
  book_val=cumunits_calc*acb,
  profit=market_val-book_val
)

## CALC PROFIT AND ROI to date - compile for currencies
curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
  mutate(currency=currency,
         roi=market_val/book_val-1)
curr_profits <- bind_rows(curr_profits, curr_profit)

```

```{r, fig.height=4, fig.width=9}
## Price v Cost
chart_title <- paste0(currency,": Price vs Cost")
pcp1 <- curr_price_cost %>% ggplot(aes(x=date, y=`BTC-CAD.Close`))+geom_line(color=price_line)+
  geom_line(aes(x=date, y=acb), color='black')+
  scale_y_continuous(labels=dollar)+
  labs(title=chart_title, x="")

## Profit
chart_title <- paste0(currency,": Net Profit")
pcp2 <- curr_price_cost %>% ggplot(aes(x=date, y=profit))+geom_col(aes(fill=profit>0))+ ## color condition
  scale_y_continuous(labels=dollar)+
  scale_fill_manual(values = setNames(c(profit,loss),c(T, F)))+ ## manual color based on condition
  theme(legend.position = "none")+ ## hide legend - not needed
  labs(title=chart_title, x="")

#grid.arrange(pcp1, pcp2, nrow=1)
## plotly version - using subplot
pcp1_p <- ggplotly(pcp1)
pcp2_p <- ggplotly(pcp2)
subplot(pcp1_p, pcp2_p, nrows=1)
```

### Ethereum

#### Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## ethereum
currency <- 'ETH'
curr_trans <- ktrans %>% filter(Currency==currency)
curr_trans$Date <- ymd_hms(curr_trans$Date) ## format timestamp
curr_trans$date_trans <- date(curr_trans$Date) ## add date field for transaction

## calc average cost (adjusted cost base)
## 2 aspects:
## - 1. basic: cumulative amount spent / cumulative units acquired (weighted average)
## - 2. account for sales: need to remove the units sold from cumulative total and 
##   reduce amount spent by number of units sold at a value of previous acb
curr_trans <- curr_trans %>% 
  mutate(
    units_calc=ifelse(Activity=='trade-buy',Amount,Amount* -1),
    cumunits_calc=cumsum(units_calc),
    cost_calc=ifelse(Activity=='trade-buy',`CAD Amount`,0),
    cost_sell_calc=ifelse(Activity=='trade-sell',units_calc*(cumsum(cost_calc)/lag(cumunits_calc)),0),
    #cumcost=ifelse(Activity=='trade-buy',cumsum(cost),abs(units)*(cumsum(cost)/cumunits)),
    #cumcostunit=ifelse(Activity=='trade-buy',cumsum(units),abs(units)),
    #cumcostunitlag=ifelse(Activity=='trade-buy',0,lag(cumunits)),
    #cumcostcost=ifelse(Activity=='trade-buy',0,cumsum(cost)),
    #cumcostcum=ifelse(Activity=='trade-buy',0,cumsum(cost)/lag(cumunits)),
    #cumcostsell=ifelse(Activity=='trade-buy',0,units*(cumsum(cost)/lag(cumunits))),
    cumcost_calc=ifelse(Activity=='trade-buy',cumsum(cost_calc)+cumsum(cost_sell_calc),cumsum(cost_calc)-abs(units_calc)*(cumsum(cost_calc)/lag(cumunits_calc))),
    acb=cumcost_calc/cumunits_calc
  )

ct1 <- curr_trans %>% ggplot(aes(x=date_trans, y=acb))+geom_col()
ct2 <- curr_trans %>% ggplot(aes(x=date_trans, y=cumunits_calc))+geom_col()
#grid.arrange(ct1, ct2, nrow=1)
## subplot instead of grid.arrange for plotly charts
subplot(ggplotly(ct1), ggplotly(ct2), nrows=1)

## SAVE (session) for currency
eth_trans_save <- curr_trans

```

```{r}
### Price Data
symbol <- "ETH-CAD" ## used in getSymbols call and chart titles
price_col <- "ETH-CAD.Close" ## needed for column references at several points below

curr <- getSymbols(Symbols=symbol, auto.assign=FALSE)
## remove today - or not, to get latest data
#curr <- curr[1:nrow(curr)-1,]

## match dates with cost data
dt_start <- as.character(min(curr_trans$date_trans))

curr_price <- curr[paste0(dt_start,"/"),price_col]

```

```{r}
## convert ta data frame and combine with cost data
curr_price_df <- as.data.frame(curr_price)
curr_price_df$date <- date(row.names(curr_price_df))
row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
curr_price_df <- curr_price_df %>% select(date, all_of(price_col)) ## all_of recommended by warnings

## combine with cost data
## get essential cols from trans 
curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
## need to have only one entry per date -> if there are multiple, should be last item for any date
curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>% summarize(cumunits_calc=last(cumunits_calc),
                                                                        cumcost_calc=last(cumcost_calc),
                                                                        acb=last(acb)
                                                                        ) %>%
  rename(date=date_trans)

## join price and cost data
curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
## fill in NA values with previous value
curr_price_cost <- na.locf(curr_price_cost)

## calc profit on any given day
## - uses !!as.name(<variable>) to identify column name based on variable set above
curr_price_cost <- curr_price_cost %>% mutate(
  market_val=!!as.name(price_col)*cumunits_calc,
  book_val=cumunits_calc*acb,
  profit=market_val-book_val
)

## calc current profits and roi - compile for currencies
curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
  mutate(currency=currency,
         roi=market_val/book_val-1)
curr_profits <- bind_rows(curr_profits, curr_profit)
```

#### Price, Costs, Profits

```{r, fig.height=4, fig.width=9}

### PRICE V COST
chart_title <- paste0(currency,": Price vs Cost")
pcp1 <- curr_price_cost %>% ggplot(aes(x=date, y=!!as.name(price_col)))+geom_line(color=price_line)+
  geom_line(aes(x=date, y=acb), color='black')+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title)

### PROFIT
chart_title <- paste0(currency,": Net Profit")
pcp2 <- curr_price_cost %>% ggplot(aes(x=date, y=profit))+geom_col(aes(fill=profit>0))+ ## color condition
  scale_y_continuous(labels=dollar)+
  scale_fill_manual(values = setNames(c(profit,loss),c(T, F)))+ ## manual color based on condition
  theme(legend.position = "none")+ ## hide legend - not needed
  labs(title=chart_title, x="")
#grid.arrange(pcp1, pcp2, nrow=1)
## plotly version - using subplot
pcp1_p <- ggplotly(pcp1)
pcp2_p <- ggplotly(pcp2)
subplot(pcp1_p, pcp2_p, nrows=1)
```

### Cardano

#### Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## cardano
currency <- 'ADA'
curr_trans <- ktrans %>% filter(Currency==currency)
curr_trans$Date <- ymd_hms(curr_trans$Date) ## format timestamp
curr_trans$date_trans <- date(curr_trans$Date) ## add date field for transaction

## calc average cost (adjusted cost base)
## 2 aspects:
## - 1. basic: cumulative amount spent / cumulative units acquired (weighted average)
## - 2. account for sales: need to remove the units sold from cumulative total and 
##   reduce amount spent by number of units sold at a value of previous acb
curr_trans <- curr_trans %>% 
  mutate(
    units_calc=ifelse(Activity=='trade-buy',Amount,Amount* -1),
    cumunits_calc=cumsum(units_calc),
    cost_calc=ifelse(Activity=='trade-buy',`CAD Amount`,0),
    cost_sell_calc=ifelse(Activity=='trade-sell',units_calc*(cumsum(cost_calc)/lag(cumunits_calc)),0),
    #cumcost=ifelse(Activity=='trade-buy',cumsum(cost),abs(units)*(cumsum(cost)/cumunits)),
    #cumcostunit=ifelse(Activity=='trade-buy',cumsum(units),abs(units)),
    #cumcostunitlag=ifelse(Activity=='trade-buy',0,lag(cumunits)),
    #cumcostcost=ifelse(Activity=='trade-buy',0,cumsum(cost)),
    #cumcostcum=ifelse(Activity=='trade-buy',0,cumsum(cost)/lag(cumunits)),
    #cumcostsell=ifelse(Activity=='trade-buy',0,units*(cumsum(cost)/lag(cumunits))),
    cumcost_calc=ifelse(Activity=='trade-buy',cumsum(cost_calc)+cumsum(cost_sell_calc),cumsum(cost_calc)-abs(units_calc)*(cumsum(cost_calc)/lag(cumunits_calc))),
    acb=cumcost_calc/cumunits_calc
  )

ct1 <- curr_trans %>% ggplot(aes(x=date_trans, y=acb))+geom_col(position=position_dodge2())
ct2 <- curr_trans %>% ggplot(aes(x=date_trans, y=cumunits_calc))+geom_col(position=position_dodge2())
#grid.arrange(ct1, ct2, nrow=1)
## subplot instead of grid.arrange for plotly charts
subplot(ggplotly(ct1), ggplotly(ct2), nrows=1)

## SAVE (session) for currency
ada_trans_save <- curr_trans
```

#### Prices, Costs, Profits

```{r}
### PRICE DATA
symbol <- "ADA-CAD" ## used in getSymbols call and chart titles
price_col <- "ADA-CAD.Close" ## needed for column references at several points below

curr <- getSymbols(Symbols=symbol, auto.assign=FALSE)
## remove today - or not if want latest price data
#curr <- curr[1:nrow(curr)-1,]

## match dates with cost data
dt_start <- as.character(min(curr_trans$date_trans))

curr_price <- curr[paste0(dt_start,"/"),price_col]
```

```{r}
## convert ta data frame and combine with cost data
curr_price_df <- as.data.frame(curr_price)
curr_price_df$date <- date(row.names(curr_price_df))
row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
curr_price_df <- curr_price_df %>% select(date, all_of(price_col)) ## all_of recommended by warnings

## combine with cost data
## get essential cols from trans 
curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
## need to have only one entry per date -> if there are multiple, should be last item for any date
curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>% summarize(cumunits_calc=last(cumunits_calc),
                                                                        cumcost_calc=last(cumcost_calc),
                                                                        acb=last(acb)
                                                                        ) %>%
  rename(date=date_trans)

## join price and cost data
curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
## fill in NA values with previous value
curr_price_cost <- na.locf(curr_price_cost)

## calc profit on any given day
## - uses !!as.name(<variable>) to identify column name based on variable set above
curr_price_cost <- curr_price_cost %>% mutate(
  market_val=!!as.name(price_col)*cumunits_calc,
  book_val=cumunits_calc*acb,
  profit=market_val-book_val
)

## calc current profits and roi - compile for currencies
curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
  mutate(currency=currency,
         roi=market_val/book_val-1)
curr_profits <- bind_rows(curr_profits, curr_profit)
```

```{r, fig.height=4, fig.width=9}
chart_title <- paste0(currency,": Price vs Cost")
pcp1 <- curr_price_cost %>% ggplot(aes(x=date, y=!!as.name(price_col)))+geom_line(color=price_line)+
  geom_line(aes(x=date, y=acb), color='black')+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title)

## PROFIT
chart_title <- paste0(currency,": Net Profit")
pcp2 <- curr_price_cost %>% ggplot(aes(x=date, y=profit))+geom_col(aes(fill=profit>0))+ ## color condition
  scale_y_continuous(labels=dollar)+
  scale_fill_manual(values = setNames(c(profit,loss),c(T, F)))+ ## manual color based on condition
  theme(legend.position = "none")+ ## hide legend - not needed
  labs(title=chart_title, x="")
  
#grid.arrange(pcp1, pcp2, nrow=1)
## plotly version - using subplot
pcp1_p <- ggplotly(pcp1)
pcp2_p <- ggplotly(pcp2)
subplot(pcp1_p, pcp2_p, nrows=1)
```

### Polkadot

#### Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## ethereum
currency <- 'DOT'
curr_trans <- ktrans %>% filter(Currency==currency)
curr_trans$Date <- ymd_hms(curr_trans$Date) ## format timestamp
curr_trans$date_trans <- date(curr_trans$Date) ## add date field for transaction

## calc average cost (adjusted cost base)
## 2 aspects:
## - 1. basic: cumulative amount spent / cumulative units acquired (weighted average)
## - 2. account for sales: need to remove the units sold from cumulative total and 
##   reduce amount spent by number of units sold at a value of previous acb
curr_trans <- curr_trans %>% 
  mutate(
    units_calc=ifelse(Activity=='trade-buy',Amount,Amount* -1),
    cumunits_calc=cumsum(units_calc),
    cost_calc=ifelse(Activity=='trade-buy',`CAD Amount`,0),
    cost_sell_calc=ifelse(Activity=='trade-sell',units_calc*(cumsum(cost_calc)/lag(cumunits_calc)),0),
    cumcost_calc=ifelse(Activity=='trade-buy',cumsum(cost_calc)+cumsum(cost_sell_calc),cumsum(cost_calc)-abs(units_calc)*(cumsum(cost_calc)/lag(cumunits_calc))),
    acb=cumcost_calc/cumunits_calc
  )

ct1 <- curr_trans %>% ggplot(aes(x=date_trans, y=acb))+geom_col(position=position_dodge2())
ct2 <- curr_trans %>% ggplot(aes(x=date_trans, y=cumunits_calc))+geom_col(position=position_dodge2())
#grid.arrange(ct1, ct2, nrow=1)
## subplot instead of grid.arrange for plotly charts
subplot(ggplotly(ct1), ggplotly(ct2), nrows=1)

## SAVE (session) for currency
ltc_trans_save <- curr_trans

```

#### Prices, Costs, Profits

```{r}
### PRICE DATA
symbol <- "DOT1-CAD" ## used in getSymbols call and chart titles
price_col <- "DOT1-CAD.Close" ## needed for column references at several points below

curr <- getSymbols(Symbols=symbol, auto.assign=FALSE)
## remove today - or not for latest price data
#curr <- curr[1:nrow(curr)-1,]

## match dates with cost data
dt_start <- as.character(min(curr_trans$date_trans))

curr_price <- curr[paste0(dt_start,"/"),price_col]
```

```{r}
## convert ta data frame and combine with cost data
curr_price_df <- as.data.frame(curr_price)
curr_price_df$date <- date(row.names(curr_price_df))
row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
curr_price_df <- curr_price_df %>% select(date, all_of(price_col)) ## all_of recommended by warnings

## combine with cost data
## get essential cols from trans 
curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
## need to have only one entry per date -> if there are multiple, should be last item for any date
curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>% summarize(cumunits_calc=last(cumunits_calc),
                                                                        cumcost_calc=last(cumcost_calc),
                                                                        acb=last(acb)
                                                                        ) %>%
  rename(date=date_trans)

## join price and cost data
curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
## fill in NA values with previous value
curr_price_cost <- na.locf(curr_price_cost)

## calc profit on any given day
## - uses !!as.name(<variable>) to identify column name based on variable set above
curr_price_cost <- curr_price_cost %>% mutate(
  market_val=!!as.name(price_col)*cumunits_calc,
  book_val=cumunits_calc*acb,
  profit=market_val-book_val
)

## calc current profits and roi - compile for currencies
curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
  mutate(currency=currency,
         roi=market_val/book_val-1)
curr_profits <- bind_rows(curr_profits, curr_profit)
```

```{r, fig.height=4, fig.width=9}
chart_title <- paste0(currency,": Price vs Cost")
pcp1 <- curr_price_cost %>% ggplot(aes(x=date, y=!!as.name(price_col)))+geom_line(color=price_line)+
  geom_line(aes(x=date, y=acb), color='black')+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title)

## PROFIT
chart_title <- paste0(currency,": Net Profit")
pcp2 <- curr_price_cost %>% ggplot(aes(x=date, y=profit))+geom_col(aes(fill=profit>0))+ ## color condition
  scale_y_continuous(labels=dollar)+
  scale_fill_manual(values = setNames(c(profit,loss),c(T, F)))+ ## manual color based on condition
  theme(legend.position = "none")+ ## hide legend - not needed
  labs(title=chart_title, x="")
  
#grid.arrange(pcp1, pcp2, nrow=1)
## plotly version - using subplot
pcp1_p <- ggplotly(pcp1)
pcp2_p <- ggplotly(pcp2)
subplot(pcp1_p, pcp2_p, nrows=1)
```

### Litecoin

#### Adjusted Cost Base (Avg Cost)

```{r, fig.height=3, fig.width=9}
## ethereum
currency <- 'LTC'
curr_trans <- ktrans %>% filter(Currency==currency)
curr_trans$Date <- ymd_hms(curr_trans$Date) ## format timestamp
curr_trans$date_trans <- date(curr_trans$Date) ## add date field for transaction

## calc average cost (adjusted cost base)
## 2 aspects:
## - 1. basic: cumulative amount spent / cumulative units acquired (weighted average)
## - 2. account for sales: need to remove the units sold from cumulative total and 
##   reduce amount spent by number of units sold at a value of previous acb
curr_trans <- curr_trans %>% 
  mutate(
    units_calc=ifelse(Activity=='trade-buy',Amount,Amount* -1),
    cumunits_calc=cumsum(units_calc),
    cost_calc=ifelse(Activity=='trade-buy',`CAD Amount`,0),
    cost_sell_calc=ifelse(Activity=='trade-sell',units_calc*(cumsum(cost_calc)/lag(cumunits_calc)),0),
    #cumcost=ifelse(Activity=='trade-buy',cumsum(cost),abs(units)*(cumsum(cost)/cumunits)),
    #cumcostunit=ifelse(Activity=='trade-buy',cumsum(units),abs(units)),
    #cumcostunitlag=ifelse(Activity=='trade-buy',0,lag(cumunits)),
    #cumcostcost=ifelse(Activity=='trade-buy',0,cumsum(cost)),
    #cumcostcum=ifelse(Activity=='trade-buy',0,cumsum(cost)/lag(cumunits)),
    #cumcostsell=ifelse(Activity=='trade-buy',0,units*(cumsum(cost)/lag(cumunits))),
    cumcost_calc=ifelse(Activity=='trade-buy',cumsum(cost_calc)+cumsum(cost_sell_calc),cumsum(cost_calc)-abs(units_calc)*(cumsum(cost_calc)/lag(cumunits_calc))),
    acb=cumcost_calc/cumunits_calc
  )

ct1 <- curr_trans %>% ggplot(aes(x=date_trans, y=acb))+geom_col()
ct2 <- curr_trans %>% ggplot(aes(x=date_trans, y=cumunits_calc))+geom_col()
#grid.arrange(ct1, ct2, nrow=1)
## subplot instead of grid.arrange for plotly charts
subplot(ggplotly(ct1), ggplotly(ct2), nrows=1)

## SAVE (session) for currency
ltc_trans_save <- curr_trans

```

#### Prices, Costs, Profits

```{r}
### PRICE DATA
symbol <- "LTC-CAD" ## used in getSymbols call and chart titles
price_col <- "LTC-CAD.Close" ## needed for column references at several points below

curr <- getSymbols(Symbols=symbol, auto.assign=FALSE)
## remove today - or not for latest price data
#curr <- curr[1:nrow(curr)-1,]

## match dates with cost data
dt_start <- as.character(min(curr_trans$date_trans))

curr_price <- curr[paste0(dt_start,"/"),price_col]
```

```{r}
## convert ta data frame and combine with cost data
curr_price_df <- as.data.frame(curr_price)
curr_price_df$date <- date(row.names(curr_price_df))
row.names(curr_price_df) <- seq(1:nrow(curr_price_df))
curr_price_df <- curr_price_df %>% select(date, all_of(price_col)) ## all_of recommended by warnings

## combine with cost data
## get essential cols from trans 
curr_trans_smry <- curr_trans %>% select(Date, date_trans, cumunits_calc, cumcost_calc, acb)
## need to have only one entry per date -> if there are multiple, should be last item for any date
curr_trans_last <- curr_trans_smry %>% group_by(date_trans) %>% summarize(cumunits_calc=last(cumunits_calc),
                                                                        cumcost_calc=last(cumcost_calc),
                                                                        acb=last(acb)
                                                                        ) %>%
  rename(date=date_trans)

## join price and cost data
curr_price_cost <- full_join(curr_price_df, curr_trans_last, by='date')
## fill in NA values with previous value
curr_price_cost <- na.locf(curr_price_cost)

## calc profit on any given day
## - uses !!as.name(<variable>) to identify column name based on variable set above
curr_price_cost <- curr_price_cost %>% mutate(
  market_val=!!as.name(price_col)*cumunits_calc,
  book_val=cumunits_calc*acb,
  profit=market_val-book_val
)

## calc current profits and roi - compile for currencies
curr_profit <- curr_price_cost %>% slice_tail() %>% select(date, market_val, book_val, profit) %>%
  mutate(currency=currency,
         roi=market_val/book_val-1)
curr_profits <- bind_rows(curr_profits, curr_profit)
```


```{r, fig.height=4, fig.width=9}
chart_title <- paste0(currency,": Price vs Cost")
pcp1 <- curr_price_cost %>% ggplot(aes(x=date, y=!!as.name(price_col)))+geom_line(color=price_line)+
  geom_line(aes(x=date, y=acb), color='black')+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title)

## PROFIT
chart_title <- paste0(currency,": Net Profit")
pcp2 <- curr_price_cost %>% ggplot(aes(x=date, y=profit))+geom_col(aes(fill=profit>0))+ ## color condition
  scale_y_continuous(labels=dollar)+
  scale_fill_manual(values = setNames(c(profit,loss),c(T, F)))+ ## manual color based on condition
  theme(legend.position = "none")+ ## hide legend - not needed
  labs(title=chart_title, x="")
  
#grid.arrange(pcp1, pcp2, nrow=1)
## plotly version - using subplot
pcp1_p <- ggplotly(pcp1)
pcp2_p <- ggplotly(pcp2)
subplot(pcp1_p, pcp2_p, nrows=1)
```

### TOTALS

```{r, fig.height=3.5, fig.width=9}
#kable(curr_profits)
curr_profits <- curr_profits %>% select(date, currency, book_val, market_val, profit, roi) %>% mutate(
  book_val=round(book_val, 0),
  market_val=round(market_val, 0),
  profit=round(profit,0),
  roi=round(roi, 2)
)

chart_title <- "Current Profit Standing by Currency"
pcp <- curr_profits %>% ggplot(aes(x=currency, y=profit))+geom_col()+
  scale_y_continuous(labels=comma)+
  labs(title=chart_title, x="")

chart_title <- "Current ROI by Currency"  
pcr <- curr_profits %>% ggplot(aes(x=currency, y=roi))+geom_col()+
  scale_y_continuous(labels=percent)+
  labs(title=chart_title, x="")

grid.arrange(pcp, pcr, nrow=1)
```

```{r}
## calc total performance
total_perform <- curr_profits %>% summarize(
  date=max(date),
  total_invested=sum(book_val),
  total_market_val=sum(market_val),
  total_profit=sum(profit),
  total_roi=sum(market_val)/sum(book_val)-1)

## functions for formattable
currency_format <- formatter("span", x ~ currency(x, digits=0))
acct_format <- formatter("span", x ~ accounting(x, digits=0, big.mark=","))
pc_format <- formatter("span", x ~ percent(x, digits=0))

formattable(curr_profits, list(
  book_val=currency_format,
  market_val=currency_format,
  profit=acct_format,
  roi=pc_format))

#curr_profits %>% kbl() %>% kable_styling()
formattable(total_perform, list(
  total_invested=currency_format,
  total_market_val=currency_format,
  total_profit=currency_format,
  total_roi=pc_format))

```

